---
title: "<span style='font-size: 70px'>Measuring and Predicting Player Fatigue in the NBA </style>"
author: "<span style='font-size: 35px'>Austin Stephen, Matthew Yep and Grace Fain </style>"
date: "<span style='font-size: 35px'> 7/13/2021 </style>"
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  cache = FALSE,
  message = FALSE,
  warning = FALSE
  )
library(readr)
library(dplyr)
library(tidyverse)
library(gt)
library(xaringanthemer)
library(teamcolors)
library(usmap)
library(ggplot2)
library(maps)
library(mapdata)
library(airball)
library(gridExtra)
library(png)
library(grid)

 
```

```{r xaringan-themer, include=FALSE, warning=FALSE }

visitors <- read_csv("./matthew_data/visitors.csv")
data1415 <- read_csv("./matthew_data/regseason1415.csv")
datos <- read_csv("./matthew_data/map1415data.csv")

style_mono_accent(base_color = "#cc002c",
  white_color = "#e6e6e6",
  background_color = "#b3b3b3",
  base_font_size = "28px"
  )
```

## Why Fatigue?

- Injury Prevention  

- Optimized Performance 

???
Very clearly established by the lit  

"Early identification and subsequent management of fatigue may prevent detrimental physical and physiological adaptations often associated with injury and enhance athletic performance and player availability"
Edwards et al The National Center for Biotechnology Information

--

```{r fig.height = 4, out.extra = 'style= "border:solid; border-color: #666666 "' }
img <- readPNG("img/fatigue.png")
 grid.raster(img)

```

???
Image courtsey of Taylor et al.

---

## Project Objectives 

1. Measure player fatigue  

???
Identify a proxy for fatigue and link it to in game performance.

--
  
2. Build a model that predict the impact of fatigue on a given 
game or multi-game sequence.  


???
This model can used by teams to examine their roster and match-ups and gain insights into 
how differnt factors are influencing their teams performance

---

## What is Fatigue

- Cannot use direct measures  

--

- Identify a proxy

???
You cannot identify a proxy for something you don't understand.

More concrete on fatigue: 
"the decline in objective performance measures derived from the capacity 
of the nervous system and contractile properties of muscles over time"
Edwards et al. The National Center for Biotechnology Information

In a perfect world we would directly measure an NBA players internal homeostasis
either with hematological techniques or with sensors that measure heart rate. However, players cannot wear heart rate monitors or tracking equipment in games and we certainly don't have data about in/post-game blood samples so we need a proxy to measure how fatigued a player was by in game events.


---

## Proxy for Fatigue

- In-game distance  

- Between game distance

???
2 major sources for measuring volume of strain that could induce player fatigue.


---

## Proxy for Fatigue   

- In-game distance  

```{r}
 tracking_player <- read_csv(
   "data/in_game_distance_data/tracking_by_player_season_daily_2018_19.csv")


 tracking_player %>%
   select(c("player_name","date","min", "dist_miles","dist_miles_off", "dist_miles_def",
            "avg_speed","avg_speed_off","avg_speed_def")) %>%
   rename("Name" = player_name,
          "Distance (mi)" = dist_miles,
          "Distance Offense"  = dist_miles_off,
          "Distance Defense" = dist_miles_def,
          "Speed (mph)" = avg_speed,
          "Speed Offense" = avg_speed_off,
          "Speed Defense" = avg_speed_def,
          "Date" = date,
          "Min" = min) %>%
   slice(1:4) %>%
   gt() %>%
   tab_header(
    title = md("**Distance By Player Individual Games**")
  ) %>%
  tab_source_note(md("Data courtesy of nba.com and Second Spectrum")) %>%
  opt_table_outline( style = "solid", width = px(5), color = "#666666")%>%
  tab_options(table.width = pct(80))
```

???
3 important things
Gen by computer vision  
Game by game  
Player level  


---

## Proxy for Fatigue

- In-game distance  

```{r}
tracking_team <- read_csv(
  "data/in_game_distance_data/tracking_by_team_season_summaries_2013_to_20.csv")

 tracking_team %>%
   select(c("team_name","w","l", "dist_miles","dist_miles_off", "dist_miles_def",
            "avg_speed","avg_speed_off","avg_speed_def")) %>%
   rename("Team" = team_name,
          "Distance (mi)" = dist_miles,
          "Distance Offense"  = dist_miles_off,
          "Distance Defense" = dist_miles_def,
          "Speed (mph)" = avg_speed,
          "Speed Offense" = avg_speed_off,
          "Speed Defense" = avg_speed_def) %>%
   slice(1:4) %>%
   gt() %>%
   tab_header(
    title = md("**Season Distance By Team**")
  ) %>%
  tab_source_note(md("Data courtesy of nba.com and Second Spectrum")) %>%
  opt_table_outline( style = "solid", width = px(5), color = "#666666") %>%
  tab_options(table.width = pct(80))
```

???
2 key differences 
Season summaries 
Team level

---

## Progress with in-game data 

- Inverse correlation between wins and in-game distance

--

```{r fig.height = 3}
team_tracking <- read_csv(
  "data/in_game_distance_data/tracking_by_team_season_summaries_2013_to_20.csv")

team_tracking %>% 
  ggplot(aes(x=dist_miles, y= w)) + 
  geom_point()+
  geom_abline(intercept = 121.6, slope = -4.73, color = "darkred")+
  theme_bw()+ 
  labs( title = "2013 to 2020 wins and distance by team",
        x = "distance (mi)",
        y = "wins")+
  theme(plot.background  = 
          element_rect(color = "#666666", size = 3, linetype ="solid" )
        ) 
```

y = 121.6 + -4.73  
p-value: .00018 and adj r-squared: .053 

???
A teams success is inversely correlated with the distance  
Opposite of what we would expect  
Causes a problem using this as proxy  

---

## Progress with in-game data 
 
- Consistent across all 7 seasons with data  

--

```{r fig.height=4}
team_tracking <- read_csv(
  "data/in_game_distance_data/tracking_by_team_season_summaries_2013_to_20.csv")

ggplot(team_tracking, aes(x = dist_miles, y= w)) +
    geom_point(size=.5, alpha = .75)+
    theme_bw()+
  facet_wrap(~season) +
  geom_smooth(method = "lm", se = FALSE)+
    labs(title = "Teams wins and average per distance per game",
         x = " distance (miles)"
         )+
  theme(plot.background  = 
          element_rect(color = "#666666", size = 3, linetype ="solid" )
        )

```

???
Convincing
Raised the question why
---

## Progress with in-game data

- At team level *not* caused by
  - 3-point FG percentage  
  - 2-point FG percentage
  - Turnovers
  - Assists 
  - Offensive rebounds 
  - Distribution of distance between bench and stars
  
???
Time I cannot go into how I showed these are not the cause.  
A lot of interesting relationships in these I don't have time to discuss

---

## Progress with in-game data 

- Teams are an abstraction

- 15 player summary statistic  

--

```{r fig.height=3}
## Low key fuck this code block 

## 2013-2020 schedule data 
schedules_2013 <- read_csv("data/schedules/season_2013_2014.csv")
schedules_2014 <- read_csv("data/schedules/season_2014_2015.csv")
schedules_2015 <- read_csv("data/schedules/season_2015_2016.csv")
schedules_2016 <- read_csv("data/schedules/season_2016_2017.csv")
schedules_2017 <- read_csv("data/schedules/season_2017_2018.csv")
schedules_2018 <- read_csv("data/schedules/season_2018_2019.csv")
schedules_2019 <- read_csv("data/schedules/season_2019_2020.csv")
schedules_2020 <- read_csv("data/schedules/season_2020_2021.csv")

schedules_2013_20 <- rbind(schedules_2013, schedules_2014, schedules_2015, 
                           schedules_2016, schedules_2017, schedules_2018, 
                           schedules_2019, schedules_2020)

## clutter the environment
rm(schedules_2013, schedules_2014, schedules_2015, 
                           schedules_2016, schedules_2017, schedules_2018, 
                           schedules_2019, schedules_2020)
## player game by game distance 14/15 season
tracking_by_player_season_daily_2014_15 <- read_csv(
  "data/in_game_distance_data/tracking_by_player_season_daily_2014_15.csv")


## player game by game box scores 14/15 season
box_game_by_game_player_2014_15 <- read_csv(
  "data/box_scores/box_scores_game_by_game_player_2014_15.csv")

tracking_by_player_season_daily_2014_15$date <- as.Date(
  tracking_by_player_season_daily_2014_15$date, "%m/%d/%Y")

# joining with schedules for game id
box_game_by_game_player_2014_15 <- merge(
    schedules_2013_20,
    box_game_by_game_player_2014_15,
    by= "idGame",
)

gbg_player_w_distance <- merge(
    tracking_by_player_season_daily_2014_15,
    box_game_by_game_player_2014_15,
    by.x= c("player_id","date"),
    by.y= c("idPlayer","dateGame")
) %>%
  select(-c("gp","namePlayer","slugTeam","idTeam","teamName","cityTeam","min1"))

## player distance 
gbg_player_w_distance <- gbg_player_w_distance %>%
  group_by(player_id) %>%
  mutate(sn_avg_dist_miles = mean(dist_miles)) %>%
  mutate(sn_avg_dist_miles_off = mean(dist_miles_off)) %>%
  mutate(sn_avg_dist_miles_def = mean(dist_miles_def)) %>%
  mutate(sn_avg_speed = mean(avg_speed)) %>%
  mutate(sn_avg_speed_off = mean(avg_speed_off)) %>%
  mutate(sn_avg_speed_def = mean(avg_speed_def)) %>%
  mutate(diff_from_sn_avg_dist = dist_miles - sn_avg_dist_miles) %>%
  mutate(diff_from_sn_avg_off = dist_miles_off - sn_avg_dist_miles_off) %>%
  mutate(diff_from_sn_avg_def = dist_miles_def - sn_avg_dist_miles_def) %>%
  mutate(diff_from_sn_avg_speed = avg_speed - sn_avg_speed) %>%
  mutate(diff_from_sn_avg_speed_def = avg_speed_def - sn_avg_speed_def) %>%
  mutate(diff_from_sn_avg_speed_off = avg_speed_off - sn_avg_speed_off)


players_5_plus_min <- gbg_player_w_distance %>% filter(min >5)


players_5_plus_min %>% 
  ggplot(aes(x=avg_speed,y =plusminus ))+
  geom_point(alpha = .05)+
  geom_smooth(method="lm")+
  labs(title = "All players with 5+ min played by game",
       caption = "plusminus is for the player",
       x = "avg speed (mph)")+
   xlim(3.25,5)+
  theme(plot.background  = 
          element_rect(color = "#666666", size = 3, linetype ="solid" )
        )
```

plusminus = 11.429 - 2.33(avg_speed)    
p-value: <2e-16

??
Really strong relationship here again. So it is not a substitution, resting  dynamic or team recruiting dynamic.
24k observations 

---

## Progress with in-game data 2

- Ideas?

???
This has me pretty stuck.

---
### Drives and schedule density data 

```{r load in data, include=FALSE}
density_drives_2014_15 <- read.csv("./data1/density_distance_drives_tracking/_density_drives_distance_daily_2014_15.csv")

density_drives_2018_19 <- read.csv("./data1/density_distance_drives_tracking/_density_drives_distance_daily_2018_19.csv")
```

```{r table}

density_drives_2014_15 %>%
   select(c("team_abbreviation","player_name","w","l", "drives", "drive_pts", "drive_pts_pct", "drive_passes_pct", "drive_tov_pct", "dist_miles", "B2B.2nd")) %>%
   rename("Team" = team_abbreviation,
          "Player" = player_name,
          "Distance (mi)" = dist_miles,
          "# of Drives" = drives,
          "2nd of a B2B" = B2B.2nd,
          "points per drive" = drive_pts_pct,
          "passes per drive" = drive_passes_pct,
          "turnovers per drive" = drive_tov_pct) %>%
   slice(5:10) %>%
   gt() %>%
   tab_header(
    title = md("**Drives, Distance, and Schedule Density by Player**")
  ) %>%
  tab_source_note(md("data courtesy of nba.com, Second Spectrum, and the airball r package")) %>%
  opt_table_outline( style = "solid", width = px(5), color = "darkgrey")

```

---
### Average Points Per Drive and Back-To-Back Games
```{r drives_pts_pct, echo=FALSE, fig.height = 5}
##ATL
b2b_2014_ATL <- density_drives_2014_15 %>%
  filter(B2B.2nd == "Yes") %>%
  filter(drives > 0) %>%
  group_by(player_id) %>%
  filter(team_abbreviation == "ATL")

other_2014_ATL <- density_drives_2014_15 %>%
  filter(B2B.2nd == "No") %>%
  filter(drives > 0) %>%
  group_by(player_id) %>%
  filter(team_abbreviation == "ATL")

notB2b_2014_ATL <- aggregate(other_2014_ATL$drive_pts_pct / nrow(other_2014_ATL), list(other_2014_ATL$player_name), FUN=sum, na.rm = TRUE)

B2b_2014_ATL <-aggregate(b2b_2014_ATL$drive_pts_pct / nrow(b2b_2014_ATL), list(b2b_2014_ATL$player_name), FUN=sum, na.rm = TRUE)

drive_pct_2014_ATL <- merge(B2b_2014_ATL, notB2b_2014_ATL, by = "Group.1") %>%
  rename("2nd_B2B_game" = x.x, "not_2nd_B2B" = x.y)

drive_2014_ATL_long <- drive_pct_2014_ATL %>%
  gather("Stat", "Value", -Group.1)

p1 <- ggplot(drive_2014_ATL_long, aes(x = Group.1, y = Value, fill = Stat)) +
  geom_bar(stat = "Identity", position = "dodge") +
  theme_bw() +
  ggtitle("Atlanta Hawks") +
  xlab("Players") +
  ylab("avg pts per drive") +
  scale_fill_manual("Stat", values = c("not_2nd_B2B" = "black", "2nd_B2B_game" = "skyblue")) +
  theme(axis.text.x = element_text(size=5, angle=45, hjust = 1))

## GSW
b2b_2014_GSW <- density_drives_2014_15 %>%
  filter(B2B.2nd == "Yes") %>%
  filter(drives > 0) %>%
  group_by(player_id) %>%
  filter(team_abbreviation == "GSW")

other_2014_GSW <- density_drives_2014_15 %>%
  filter(B2B.2nd == "No") %>%
  filter(drives > 0) %>%
  group_by(player_id) %>%
  filter(team_abbreviation == "GSW")

notB2b_2014_GSW <- aggregate(other_2014_GSW$drive_pts_pct / nrow(other_2014_GSW), list(other_2014_GSW$player_name), FUN=sum, na.rm = TRUE)

b2b_2014_GSW <-aggregate(b2b_2014_GSW$drive_pts_pct / nrow(b2b_2014_GSW), list(b2b_2014_GSW$player_name), FUN=sum, na.rm = TRUE)

drive_pct_2014_GSW <- merge(b2b_2014_GSW, notB2b_2014_GSW, by = "Group.1") %>%
  rename("2nd_B2B_game" = x.x, "not_2nd_B2B" = x.y)

drive_2014_GSW_long <- drive_pct_2014_GSW %>%
  gather("Stat", "Value", -Group.1)

p2 <- ggplot(drive_2014_GSW_long, aes(x = Group.1, y = Value, fill = Stat)) +
  geom_bar(stat = "Identity", position = "dodge") +
  theme_bw() +
  ggtitle("Golden State Warriors") +
  xlab("Players") +
  ylab("avg pts per drive") +
  scale_fill_manual("Stat", values = c("not_2nd_B2B" = "black", "2nd_B2B_game" = "skyblue")) +
  theme(axis.text.x = element_text(size=5, angle=45, hjust = 1))

##MIN
b2b_2014_MIN <- density_drives_2014_15 %>%
  filter(B2B.2nd == "Yes") %>%
  filter(drives > 0) %>%
  group_by(player_id) %>%
  filter(team_abbreviation == "MIN")

other_2014_MIN <- density_drives_2014_15 %>%
  filter(B2B.2nd == "No") %>%
  filter(drives > 0) %>%
  group_by(player_id) %>%
  filter(team_abbreviation == "MIN")

notB2b_2014_MIN <- aggregate(other_2014_MIN$drive_pts_pct / nrow(other_2014_MIN), list(other_2014_MIN$player_name), FUN=sum, na.rm = TRUE)

b2b_2014_MIN <-aggregate(b2b_2014_MIN$drive_pts_pct / nrow(b2b_2014_MIN), list(b2b_2014_MIN$player_name), FUN=sum, na.rm = TRUE)

drive_pct_2014_MIN <- merge(b2b_2014_MIN, notB2b_2014_MIN, by = "Group.1") %>%
  rename("2nd_B2B_game" = x.x, "not_2nd_B2B" = x.y)

drive_2014_MIN_long <- drive_pct_2014_MIN %>%
  gather("Stat", "Value", -Group.1)

p3 <- ggplot(drive_2014_MIN_long, aes(x = Group.1, y = Value, fill = Stat)) +
  geom_bar(stat = "Identity", position = "dodge") +
  theme_bw() +
  ggtitle("Minnesota Timberwolves") +
  xlab("Players") +
  ylab("avg pts per drive") +
  scale_fill_manual("Stat", values = c("not_2nd_B2B" = "black", "2nd_B2B_game" = "skyblue")) +
  theme(axis.text.x = element_text(size=5, angle=45, hjust = 1))

## NYK
b2b_2014_NYK <- density_drives_2014_15 %>%
  filter(B2B.2nd == "Yes") %>%
  filter(drives > 0) %>%
  group_by(player_id) %>%
  filter(team_abbreviation == "NYK") %>%
  arrange(drive_pts_pct)

other_2014_NYK <- density_drives_2014_15 %>%
  filter(B2B.2nd == "No") %>%
  filter(drives > 0) %>%
  group_by(player_id) %>%
  filter(team_abbreviation == "NYK") %>%
  arrange(drive_pts_pct)

notB2b_2014_NYK <- aggregate(other_2014_NYK$drive_pts_pct / nrow(other_2014_NYK), list(other_2014_NYK$player_name), FUN=sum, na.rm = TRUE)

b2b_2014_NYK <-aggregate(b2b_2014_NYK$drive_pts_pct / nrow(b2b_2014_NYK), list(b2b_2014_NYK$player_name), FUN=sum, na.rm = TRUE)

drive_pct_2014_NYK <- merge(b2b_2014_NYK, notB2b_2014_NYK, by = "Group.1") %>%
  rename("2nd_B2B_game" = x.x, "not_2nd_B2B" = x.y)

drive_2014_NYK_long <- drive_pct_2014_NYK %>%
  gather("Stat", "Value", -Group.1)
  
p4 <- ggplot(drive_2014_NYK_long, aes(x = Group.1, y = Value, fill = Stat)) +
  geom_bar(stat = "Identity", position = "dodge") +
  theme_bw()  +
  ggtitle("New York Knicks") +
  xlab("Players") +
  ylab("avg pts per drive") +
  scale_fill_manual("Stat", values = c("not_2nd_B2B" = "black", "2nd_B2B_game" = "skyblue")) +
  theme(axis.text.x = element_text(size=5, angle=45, hjust = 1))

grid.arrange(p2, p3, p1, p4, nrow = 2, ncol = 2)
```

---
### Over half of the NBA teams score less points per drive in the second game of a back-to-back
```{r team pts pct, echo = FALSE, fig.height =5}

##ALL
b2b_2014_ALL <- density_drives_2014_15 %>%
  filter(B2B.2nd == "Yes") %>%
  filter(drives > 0) %>%
  group_by(team_abbreviation) 

other_2014_ALL <- density_drives_2014_15 %>%
  filter(B2B.2nd == "No") %>%
  filter(drives > 0) %>%
  group_by(team_abbreviation) 

notB2b_2014_ALL_pts <- aggregate(other_2014_ALL$drive_pts_pct / nrow(other_2014_ALL), list(other_2014_ALL$team_abbreviation), FUN=sum, na.rm = TRUE)

B2b_2014_ALL_pts <-aggregate(b2b_2014_ALL$drive_pts_pct / nrow(b2b_2014_ALL), list(b2b_2014_ALL$team_abbreviation), FUN=sum, na.rm = TRUE)

drive_pct_2014_ALL_pts <- merge(B2b_2014_ALL_pts, notB2b_2014_ALL_pts, by = "Group.1") %>%
  rename("2nd_B2B_game" = x.x, "not_2nd_B2B" = x.y)

drive_2014_ALL_long_pts <- drive_pct_2014_ALL_pts %>%
  gather("Stat", "Value", -Group.1) 

ggplot(drive_2014_ALL_long_pts, aes(x = Group.1, y = Value, fill = Stat)) +
  geom_bar(stat = "Identity", position = "dodge") +
  theme_bw() +
  ggtitle("Average points per drive for each NBA team") +
  xlab("Team") +
  ylab("avg points per drive") +
  scale_fill_manual("Stat", values = c("not_2nd_B2B" = "black", "2nd_B2B_game" = "skyblue")) +
  theme(axis.text.x = element_text(size=7, angle=45, hjust = 1))



```
---

### Average number of passes out of drives and back-to-back games
```{r drive and pass, echo = FALSE, fig.height = 5}
##ATL
b2b_2014_ATL <- density_drives_2014_15 %>%
  filter(B2B.2nd == "Yes") %>%
  filter(drives > 0) %>%
  group_by(player_id) %>%
  filter(team_abbreviation == "ATL")

other_2014_ATL <- density_drives_2014_15 %>%
  filter(B2B.2nd == "No") %>%
  filter(drives > 0) %>%
  group_by(player_id) %>%
  filter(team_abbreviation == "ATL")

notB2b_2014_ATL_pass <- aggregate(other_2014_ATL$drive_passes_pct / nrow(other_2014_ATL), list(other_2014_ATL$player_name), FUN=sum, na.rm = TRUE)

B2b_2014_ATL_pass <-aggregate(b2b_2014_ATL$drive_passes_pct / nrow(b2b_2014_ATL), list(b2b_2014_ATL$player_name), FUN=sum, na.rm = TRUE)

drive_pct_2014_ATL_pass <- merge(B2b_2014_ATL_pass, notB2b_2014_ATL_pass, by = "Group.1") %>%
  rename("2nd_B2B_game" = x.x, "not_2nd_B2B" = x.y)

drive_2014_ATL_long_pass <- drive_pct_2014_ATL_pass %>%
  gather("Stat", "Value", -Group.1)

p1 <- ggplot(drive_2014_ATL_long_pass, aes(x = Group.1, y = Value, fill = Stat)) +
  geom_bar(stat = "Identity", position = "dodge") +
  theme_bw() +
  ggtitle("Atlanta Hawks") +
  xlab("Players") +
  ylab("avg passes per drive") +
  scale_fill_manual("Stat", values = c("not_2nd_B2B" = "black", "2nd_B2B_game" = "skyblue")) +
  theme(axis.text.x = element_text(size=5, angle=45, hjust = 1))

## GSW
b2b_2014_GSW <- density_drives_2014_15 %>%
  filter(B2B.2nd == "Yes") %>%
  filter(drives > 0) %>%
  group_by(player_id) %>%
  filter(team_abbreviation == "GSW")

other_2014_GSW <- density_drives_2014_15 %>%
  filter(B2B.2nd == "No") %>%
  filter(drives > 0) %>%
  group_by(player_id) %>%
  filter(team_abbreviation == "GSW")

notB2b_2014_GSW_pass <- aggregate(other_2014_GSW$drive_passes_pct / nrow(other_2014_GSW), list(other_2014_GSW$player_name), FUN=sum, na.rm = TRUE)

b2b_2014_GSW_pass <-aggregate(b2b_2014_GSW$drive_passes_pct/ nrow(b2b_2014_GSW), list(b2b_2014_GSW$player_name), FUN=sum, na.rm = TRUE)

drive_pct_2014_GSW_pass <- merge(b2b_2014_GSW_pass, notB2b_2014_GSW_pass, by = "Group.1") %>%
  rename("2nd_B2B_game" = x.x, "not_2nd_B2B" = x.y)

drive_2014_GSW_long_pass <- drive_pct_2014_GSW_pass %>%
  gather("Stat", "Value", -Group.1)

p2 <- ggplot(drive_2014_GSW_long_pass, aes(x = Group.1, y = Value, fill = Stat)) +
  geom_bar(stat = "Identity", position = "dodge") +
  theme_bw() +
  ggtitle("Golden State Warriors") +
  xlab("Players") +
  ylab("avg passes per drive") +
  scale_fill_manual("Stat", values = c("not_2nd_B2B" = "black", "2nd_B2B_game" = "skyblue")) +
  theme(axis.text.x = element_text(size=5, angle=45, hjust = 1))

##MIN
b2b_2014_MIN <- density_drives_2014_15 %>%
  filter(B2B.2nd == "Yes") %>%
  filter(drives > 0) %>%
  group_by(player_id) %>%
  filter(team_abbreviation == "MIN")

other_2014_MIN <- density_drives_2014_15 %>%
  filter(B2B.2nd == "No") %>%
  filter(drives > 0) %>%
  group_by(player_id) %>%
  filter(team_abbreviation == "MIN")

notB2b_2014_MIN_pass <- aggregate(other_2014_MIN$drive_passes_pct / nrow(other_2014_MIN), list(other_2014_MIN$player_name), FUN=sum, na.rm = TRUE)

b2b_2014_MIN_pass <-aggregate(b2b_2014_MIN$drive_passes_pct / nrow(b2b_2014_MIN), list(b2b_2014_MIN$player_name), FUN=sum, na.rm = TRUE)

drive_pct_2014_MIN_pass <- merge(b2b_2014_MIN_pass, notB2b_2014_MIN_pass, by = "Group.1") %>%
  rename("2nd_B2B_game" = x.x, "not_2nd_B2B" = x.y)

drive_2014_MIN_long_pass <- drive_pct_2014_MIN_pass %>%
  gather("Stat", "Value", -Group.1)

p3 <- ggplot(drive_2014_MIN_long_pass, aes(x = Group.1, y = Value, fill = Stat)) +
  geom_bar(stat = "Identity", position = "dodge") +
  theme_bw() +
  ggtitle("Minnesota Timberwolves") +
  xlab("Players") +
  ylab("avg passes per drive") +
  scale_fill_manual("Stat", values = c("not_2nd_B2B" = "black", "2nd_B2B_game" = "skyblue")) +
  theme(axis.text.x = element_text(size=5, angle=45, hjust = 1))

## NYK
b2b_2014_NYK <- density_drives_2014_15 %>%
  filter(B2B.2nd == "Yes") %>%
  filter(drives > 0) %>%
  group_by(player_id) %>%
  filter(team_abbreviation == "NYK") 

other_2014_NYK <- density_drives_2014_15 %>%
  filter(B2B.2nd == "No") %>%
  filter(drives > 0) %>%
  group_by(player_id) %>%
  filter(team_abbreviation == "NYK") 

notB2b_2014_NYK_pass <- aggregate((other_2014_NYK$drive_passes_pct) / nrow(other_2014_NYK), list(other_2014_NYK$player_name), FUN=sum, na.rm = TRUE)

b2b_2014_NYK_pass <-aggregate((b2b_2014_NYK$drive_passes_pct) / nrow(b2b_2014_NYK), list(b2b_2014_NYK$player_name), FUN=sum, na.rm = TRUE)

drive_pct_2014_NYK_pass <- merge(b2b_2014_NYK_pass, notB2b_2014_NYK_pass, by = "Group.1") %>%
  rename("2nd_B2B_game" = x.x, "not_2nd_B2B" = x.y)

drive_2014_NYK_long_pass <- drive_pct_2014_NYK_pass %>%
  gather("Stat", "Value", -Group.1)
  
p4 <- ggplot(drive_2014_NYK_long_pass, aes(x = Group.1, y = Value, fill = Stat)) +
  geom_bar(stat = "Identity", position = "dodge") +
  theme_bw()  +
  ggtitle("New York Knicks") +
  xlab("Players") +
  ylab("avg passes per drive") +
  scale_fill_manual("Stat", values = c("not_2nd_B2B" = "black", "2nd_B2B_game" = "skyblue")) +
  theme(axis.text.x = element_text(size=5, angle=45, hjust = 1))


grid.arrange(p2, p3, p1, p4, nrow = 2, ncol = 2)

```

---
### Over half of the NBA teams pass out of drives more in the second game of a back-to-back
```{r 2018-2019 season, echo = FALSE, fig.height =5}

##ALL
b2b_2014_ALL <- density_drives_2014_15 %>%
  filter(B2B.2nd == "Yes") %>%
  filter(drives > 0) %>%
  group_by(team_abbreviation) 

other_2014_ALL <- density_drives_2014_15 %>%
  filter(B2B.2nd == "No") %>%
  filter(drives > 0) %>%
  group_by(team_abbreviation) 

notB2b_2014_ALL_pass <- aggregate(other_2014_ALL$drive_passes_pct / nrow(other_2014_ALL), list(other_2014_ALL$team_abbreviation), FUN=sum, na.rm = TRUE)

B2b_2014_ALL_pass <-aggregate(b2b_2014_ALL$drive_passes_pct / nrow(b2b_2014_ALL), list(b2b_2014_ALL$team_abbreviation), FUN=sum, na.rm = TRUE)

drive_pct_2014_ALL_pass <- merge(B2b_2014_ALL_pass, notB2b_2014_ALL_pass, by = "Group.1") %>%
  rename("2nd_B2B_game" = x.x, "not_2nd_B2B" = x.y)

drive_2014_ALL_long_pass <- drive_pct_2014_ALL_pass %>%
  gather("Stat", "Value", -Group.1) 

ggplot(drive_2014_ALL_long_pass, aes(x = Group.1, y = Value, fill = Stat)) +
  geom_bar(stat = "Identity", position = "dodge") +
  theme_bw() +
  ggtitle("Average percent of drives that are passed by each NBA team") +
  xlab("Team") +
  ylab("avg # of passes per drive") +
  scale_fill_manual("Stat", values = c("not_2nd_B2B" = "black", "2nd_B2B_game" = "skyblue")) +
  theme(axis.text.x = element_text(size=7, angle=45, hjust = 1))


```

---
### Less than half of the NBA teams turnover the ball on drives more in the second game of a back-to-back game
```{r drives tov, echo = FALSE, fig.height = 4}
##ALL
b2b_2014_ALL <- density_drives_2014_15 %>%
  filter(B2B.2nd == "Yes") %>%
  filter(drives > 0) %>%
  group_by(team_abbreviation) 

other_2014_ALL <- density_drives_2014_15 %>%
  filter(B2B.2nd == "No") %>%
  filter(drives > 0) %>%
  group_by(team_abbreviation) 

notB2b_2014_ALL_tov <- aggregate(other_2014_ALL$drive_tov_pct / nrow(other_2014_ALL), list(other_2014_ALL$team_abbreviation), FUN=sum, na.rm = TRUE)

B2b_2014_ALL_tov <-aggregate(b2b_2014_ALL$drive_tov_pct / nrow(b2b_2014_ALL), list(b2b_2014_ALL$team_abbreviation), FUN=sum, na.rm = TRUE)

drive_pct_2014_ALL_tov <- merge(B2b_2014_ALL_tov, notB2b_2014_ALL_tov, by = "Group.1") %>%
  rename("2nd_B2B_game" = x.x, "not_2nd_B2B" = x.y)

drive_2014_ALL_long_tov <- drive_pct_2014_ALL_tov %>%
  gather("Stat", "Value", -Group.1) 

ggplot(drive_2014_ALL_long_tov, aes(x = Group.1, y = Value, fill = Stat)) +
  geom_bar(stat = "Identity", position = "dodge") +
  theme_bw() +
  ggtitle("Average percent of drives that are turned-over by each NBA team") +
  xlab("Team") +
  ylab("avg percent of turnovers per drive") +
  scale_fill_manual("Stat", values = c("not_2nd_B2B" = "black", "2nd_B2B_game" = "skyblue")) +
  theme(axis.text.x = element_text(size=7, angle=45, hjust = 1))
```

---
### How does travel between games affect team performance?
```{r Map, echo = FALSE, fig.height= 6.5, fig.width=12}
cumulative_distance1415 <- data1415 %>%
  group_by(Team) %>%
  summarise_at(vars(Distance),
               list(total_distance_traveled = sum))

home_stadium_coords <- data1415 %>%
  filter(Visitor == FALSE) %>%
  select(Team, Latitude, Longitude) %>%
  distinct()

coords_and_distance <- merge(x = cumulative_distance1415,  y= home_stadium_coords,
                             by = "Team")

highlight <- coords_and_distance %>%
  filter(Team %in% c("Portland Trail Blazers","Denver Nuggets", 
                     "Cleveland Cavaliers", "Boston Celtics"))

usa <- map_data("usa")
states <- map_data("state")
ggplot() + geom_polygon(data = states, aes(x=long, y = lat, fill = region, group = group), fill = "azure2", color = "gray") + 
  coord_fixed(1.3) +
  geom_point(data = coords_and_distance, aes(x=Longitude, y=Latitude,  
                                             size = sqrt(total_distance_traveled)),
             alpha = 0.8) + 
  theme(axis.line=element_blank(),axis.text.x=element_blank(),
        axis.text.y=element_blank(),axis.ticks=element_blank(),
        axis.title.x=element_blank(),
        axis.title.y=element_blank(),
        panel.background=element_blank(),panel.border=element_blank(),panel.grid.major=element_blank(),
        panel.grid.minor=element_blank(),plot.background=element_blank(),
        legend.position = "top") +
  geom_point(data = highlight, aes(x=Longitude, y=Latitude,colour = "red", size = sqrt(total_distance_traveled))) +
  geom_text(data = coords_and_distance, aes(x=Longitude, y=Latitude, 
                                            label=ifelse(Team %in% c("Portland Trail Blazers","Denver Nuggets", 
                                         "Cleveland Cavaliers", "Boston Celtics"),
                             as.character(Team),'')), hjust= 0,vjust=0)
```

---

### The Cavaliers also cross far less timezones than the coastal teams
```{r Flight Paths, echo = FALSE, fig.height = 7, fig.width=12}
nba_travel_plot(data = datos,
                season = 2015,
                team = c("Portland Trail Blazers","Denver Nuggets", "Cleveland Cavaliers", "Boston Celtics"),
                city_color = "yellow",
                plot_background_fill = "black",
                land_color = "gray",
                caption_color = "lightblue",
                ncolumns = 2)
```

---

### Travel, Schedule and Density Data

```{r Matthew data, echo=FALSE, fig.align = "center", fig.height = 16,fig.width = 40}
img <- readPNG("./images/example_data.png") 
grid.raster(img)
```
- **Granularity:** Each row represents a game. The stats are relative to the visiting team in the *Team* column
---

### Linear Model: Distance doesn't have a significant effect on the score diff
score_diff = $\beta_0$ + $\beta_1$ win_percent_diff + $\beta_2$distance + $\beta_3$rest + $\beta_4$b2b_2nd + $\beta_5$three_in_four + $\beta$'s for each shift

```{r Linear Model, echo=FALSE, fig.align = "center", fig.height = 4.5, fig.width = 6}
img2 <- readPNG("./images/summary_visitors_model.png") 
grid.raster(img2)
```
---

#What's on deck
- Find stronger proxies to account for difference in team vs opponent strength
- Logistic regression model predicting win or loss instead of score difference
- Test the models on Bubble and 2021 data (where there was none of significantly less travel/shift between timezones)

