---
title: "<span style='font-size: 70px'>Measuring and Predicting Player Fatigue in the NBA </style>"
author: "<span style='font-size: 35px'>Austin Stephen, Matthew Yep and Grace Fain </style>"
date: "<span style='font-size: 35px'> 7/13/2021 </style>"
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  cache = FALSE,
  message = FALSE,
  warning = FALSE
  )
library(readr)
library(dplyr)
library(tidyverse)
library(gt)
library(xaringanthemer)
library(teamcolors)
library(usmap)
library(ggplot2)
library(maps)
library(mapdata)
library(airball)
library(gridExtra)
library(png)
library(grid)

 
```

```{r xaringan-themer, include=FALSE, warning=FALSE }

visitors <- read_csv("./matthew_data/visitors.csv")
data1415 <- read_csv("./matthew_data/regseason1415.csv")
datos <- read_csv("./matthew_data/map1415data.csv")

style_mono_accent(base_color = "#cc002c",
  white_color = "#e6e6e6",
  background_color = "#b3b3b3",
  base_font_size = "28px"
  )
```


## Why Fatigue?

- Injury Prevention  

- Optimized Performance 

???
Very clearly established by the lit  

"Early identification and subsequent management of fatigue may prevent detrimental physical and physiological adaptations often associated with injury and enhance athletic performance and player availability"
Edwards et al The National Center for Biotechnology Information

--

```{r fig.height = 4, out.extra = 'style= "border:solid; border-color: #666666 "' }
img <- readPNG("img/fatigue.png")
 grid.raster(img)

```

???
Image courtsey of Taylor et al.

---

## Project Objectives 

1. Measure player fatigue  

???
Identify a proxy for fatigue and link it to in game performance.

--
  
2. Build a model that predict the impact of fatigue on a given 
game or multi-game sequence.  


???
This model can used by teams to examine their roster and match-ups and gain insights into 
how differnt factors are influencing their teams performance

---


