---
title: "EDA for NBA"
author: "Austin Stephen"
date: "6/23/2021"
output:
  pdf_document: default
  html_document: default
---
```{r setup, include=FALSE }
library(tidyverse)
library(dplyr)
library(nbastatR)
library(NBAr)
knitr::opts_chunk$set(
  cache = TRUE,
  echo = FALSE,
  message = FALSE, 
  warning = FALSE,
  fig.height=3
)

## Setting up the importing datasets 
tracking_players <- read.csv(
  "../data/in_game_distance_data/tracking_by_player_season_summaries_2013_to_20.csv")

tracking_teams <- read.csv(
  "../data/in_game_distance_data/tracking_by_team_season_summaries_2013_to_20.csv")

players_dictonary <- read.csv("../data/player_tables/player_table_reduced_2010_to_21.csv")

player_stats <- read.csv(
  "../data/in_game_distance_data/tracking_by_player_season_summaries_2013_to_20.csv")

## adding player position to tracking players
tracking_players <-  left_join(tracking_players,
                               players_dictonary,
                               by= c("player_id" = "idPlayer"))

```
## Distance Traveled in Game 

Take Aways:   

* Players cover more distance on offense than defense.This makes sense, running plays requires more movement than defending them. 
* Over the last 7 seasons distance is roughly similar  
* Guards move the most, forwards the second most, and centers the least. This
spread between positions is approximately the same across offense and defense
*  There may be an inverse correlation between distance a team travels per game and wins.
Are they more efficient? Is this not the opposite of what we expect?

#### Distribution of Distance  

```{r, fig.height=3}
## Game Distance Empirical CDFs
ggplot(tracking_players,aes(x = dist_miles)) + 
  stat_ecdf() + 
  theme_bw() +
  geom_rug(alpha = .005)+
  geom_abline( intercept = 0.5, slope = 0, color = "red")+
  labs(title = "Distance Traveled in Game for Players CDF",
       x = "Distance (miles)")

ggplot(tracking_players) + 
  stat_ecdf(aes(x = dist_miles_off, color = "offense")) + 
  stat_ecdf(aes(x = dist_miles_def, color = "defense"))+
  theme_bw() +
  geom_rug(alpha = .005)+
  geom_abline( intercept = 0.5, slope = 0, color = "red",)+
  labs(title = "Distance Offense v. Defense CDF",
       x = "Distance (miles)", colour = "Type")+
  theme(legend.position = "right")


```

__Entire Game:__ min is 0.020, max is 2.990, and mean is 1.425 miles  
__Offense:__ min is 0.020, max is 1.560, and mean is .765 miles  
__Defense:__ min is .010, max is 1.29, and mean is .671 miles  

### Season to Season  

```{r}
year_2013 <- tracking_players %>% filter(season  == 2013) 
year_2014 <- tracking_players %>% filter(season  == 2014)
year_2015 <- tracking_players %>% filter(season  == 2015)
year_2016 <- tracking_players %>% filter(season  == 2016)
year_2017 <- tracking_players %>% filter(season  == 2017)
year_2018 <- tracking_players %>% filter(season  == 2018)

ggplot(tracking_players, size = .1, aes(x = dist_miles), alpha = .5) + 
  stat_ecdf(data = year_2013, aes(color = "2013")) + 
  stat_ecdf(data = year_2014, aes(color = "2014")) +
  stat_ecdf(data = year_2015, aes(color = "2015")) +
  stat_ecdf(data = year_2016, aes(color = "2016")) +
  stat_ecdf(data = year_2017, aes(color = "2017")) +
  stat_ecdf(data = year_2018, aes(color = "2018")) +
  theme_bw() +
  geom_rug(alpha = .005)+
  geom_abline( intercept = 0.5, slope = 0, color = "red",)+
  labs(title = "Player Distance by year CDF",
       x = "Distance(miles) ", color = "Year")
```
```{r}

```





### Player Position  

```{r}
guards <- tracking_players %>% filter(PlayerPosition == "G")
forwards <- tracking_players %>% filter(PlayerPosition == "F")
centers <- tracking_players %>% filter(PlayerPosition == "C")

ggplot(tracking_players, aes(x = dist_miles), color = PlayerPosition) +
  stat_ecdf(data = guards,aes( color = "Guard")) +
  stat_ecdf(data = forwards,aes( color = "Forward")) +
  stat_ecdf(data = centers, aes(color = "Center")) +
  theme_bw() +
  geom_rug(alpha = .005)+
  geom_abline( intercept = 0.5, slope = 0, color = "red",)+
  labs(title = "Distance Traveled in Game By Position CDF",
       x = "Distance(miles)", color = "Position")

ggplot(tracking_players, aes(x = dist_miles_off), color = PlayerPosition) +
  stat_ecdf(data = guards,aes( color = "Guard")) +
  stat_ecdf(data = forwards,aes( color = "Forward")) +
  stat_ecdf(data = centers, aes(color = "Center")) +
  theme_bw() +
  geom_rug(alpha = .005)+
  geom_abline( intercept = 0.5, slope = 0, color = "darkred",)+
  labs(title = "Distance on Offense By Position CDF",
       x = "Distance(miles)", color = "Position")

ggplot(tracking_players, aes(x = dist_miles_def), color = PlayerPosition) +
  stat_ecdf(data = guards,aes( color = "Guard")) +
  stat_ecdf(data = forwards,aes( color = "Forward")) +
  stat_ecdf(data = centers, aes(color = "Center")) +
  theme_bw() +
  geom_rug(alpha = .005)+
  geom_abline( intercept = 0.5, slope = 0, color = "darkred",)+
  labs(title = "Distance Defense By Position CDF",
       x = "Distance(miles)", color = "Position")
```

### Speed in Game 
```{r}
tracking_players %>%
  filter(avg_speed >3 & avg_speed < 5.5) %>%
ggplot(aes(x = avg_speed)) + 
  stat_ecdf() + 
  theme_bw() +
  geom_rug(alpha = .005)+
  geom_abline( intercept = 0.5, slope = 0, color = "red")+
  labs(title = "Player Average Speed CDF")

```

### Scatter plots looking for relationship with distance
```{r}
ggplot(tracking_teams, aes(x = dist_miles, y= w)) +
    geom_point(size=.5, alpha = .75)+
    theme_bw()+
  facet_wrap(~season) +
    labs(title = "Teams wins and average per distance per game",
         x = " distance (miles)",
         caption= "**distance is the sum of players on the team per game averaged")
```

```{r}
## are the best players at a position moving more?
## does shooting percentage go down as distance increases?
## Need to find quarter or minute distance data. 
## is distance covered on defense a predictor of rebounds?

```

