---
html_document: default
author: "Austin Stephen"
date: "7/13/2021"
output:
  html_document: default
title:  "NBA Fatigue"
---
  
```{r setup, include=FALSE }
library(nbastatR)
library(NBAr)
library(ballr)
library(tidyverse)
library(dplyr)
knitr::opts_chunk$set(
  cache = TRUE,
  echo = FALSE,
  message = FALSE, 
  warning = FALSE,
  fig.height=3
)

## player game by game distance 14/15 season
tracking_by_player_season_daily_2014_15 <- read.csv(
  "../data/in_game_distance_data/tracking_by_player_season_daily_2014_15.csv")


## player game by game box scores 14/15 season
box_game_by_game_player_2014_15 <- read.csv(
  "../data/box_scores/box_scores_game_by_game_player_2014_15.csv")


## Player id and position map table 
player_table <- read.csv(
  "../data/player_tables/player_table_all_2010_to_21.csv")
  
## player tracking distance 
player_tracking <- read.csv(
  "../data/in_game_distance_data/tracking_by_player_season_summaries_2013_to_20.csv")

## team tracking distance 
team_tracking <- read.csv(
  "../data/in_game_distance_data/tracking_by_team_season_summaries_2013_to_20.csv")

## Player season box scores
season_box_scores_player_2010_20 <- read.csv(
  "../data/box_scores/season_box_scores_player_2010_20.csv")

## Team season box scores
season_box_scores_team_2010_20 <- read.csv(
  "../data/box_scores/season_box_scores_team_2010_20.csv")

## game by game player distance joined to player box scores.
gbg_player_w_distance <- read_csv(
 "../data/in_game_distance_data/in_game_dist_box_join.csv")

## windows of time analysis 
layer_5dayWindow_dist_daily_2014_15 <- read_csv(
  "../data/in_game_distance_data/player_5dayWindow_dist_daily_2014_15.csv") %>%
  rename(wind_5_dist = "dist_5dayWindow") %>% select(player_id, date,wind_5_dist)

layer_5dayWindow_dist_daily_2014_15$date <- as.Date(
  layer_5dayWindow_dist_daily_2014_15$date, "%m/%d/%Y")

## removing all instances where player played less than 15 seconds because 
## these are just not insightful instances 
gbg_player_w_distance <- gbg_player_w_distance %>% filter(min > .25)

gbg_player_w_distance_5_day_window <- merge(x = gbg_player_w_distance,
                               y = layer_5dayWindow_dist_daily_2014_15,
                               by.x =  c("player_id","date" ),
                               by.y = c("player_id","date" )) %>%
  filter(wind_5_dist != 9999) 

## Adding a difference from mean 5 day window
gbg_player_w_distance_5_day_window <- gbg_player_w_distance_5_day_window %>%
  group_by(player_id) %>%
  mutate(sn_avg_wind_5_dist = mean(wind_5_dist) ) %>%
  mutate(diff_sn_avg_wind_5_dist = wind_5_dist - sn_avg_wind_5_dist )

## Distance differential at a given game

## Cross with Matthew's distance data to see if there is a correlation between the distance 
#  outside of a game and distance in game.

## See if teams adapt their styles to their opponents. So if defense distance is 
# correlated with opponent offense by season average and game.

## see if speed or distance differential is related to score score differential 

## Scatter plot by team that includes player name showing if having 1 outlier player 
#  leads to a decrease in the overall distance traveled.

```

## RoadMap of Goals:
1. Control for enough factors to suggest a relationship between in-game distance 
and athlete performance. (completed) 
2. Find evidence(or lack thereof) of player fatigue in a multi-game sequence 
and the expected decline in performance follows. (in-progress)   
   * (assuming success in first step) Identify the threshold where I can no 
longer observe a decline in performance suggesting this is the optimal between 
session recovery interval. (in-progress)   

## 1.  

### In-game Prefromance and Distance

#### Distance 

**Hypothesis:** More distance is linked to improved performance metrics 
because it serves as a proxy for athletes exertion in the game.   

**Why it matters:** If we offer evidence distance has a relationship with
player performance then we are justified in exploring factors that inhibit a 
players ability to cover distance, like  multi-game fatigue.

**Result:** Evidence to support the hypothesis is found when controlling for the 
players baseline performance level.

**Note on the data** All instances where a player played less than 15 seconds 
in the whole game were removed, for 3pt percent one three must have been 
attempted, for fg percent 1 fg must have been attempted.  

```{r include = FALSE}
##fgpct all distance
gbg_player_w_distance_at <-  gbg_player_w_distance %>%
  filter(fga > 0 )

gbg_player_w_distance %>%
ggplot(aes(x=diff_from_sn_avg_fgpct, y =diff_from_sn_avg_dist ))+
  geom_point(alpha = .05)+
  geom_smooth(method="lm")

model <- lm(diff_from_sn_avg_fgpct ~ diff_from_sn_avg_dist, 
            data = gbg_player_w_distance_at )
summary(model)

# fg3pct all distance
gbg_player_w_distance_at3 <-  gbg_player_w_distance %>%
  filter(fg3a > 0 )

gbg_player_w_distance_at3 %>%
ggplot(aes(x=diff_from_sn_avg_fg3pct, y =diff_from_sn_avg_dist ))+
  geom_point(alpha = .05)+
  geom_smooth(method="lm")

model <- lm(diff_from_sn_avg_fg3pct ~ diff_from_sn_avg_dist, 
            data = gbg_player_w_distance_at3 )
summary(model)

# ftpct all distance
gbg_player_w_distance %>%
  filter(!is.na(ftpct)) %>%
ggplot(aes(x=diff_from_sn_avg_ftpct, y =diff_from_sn_avg_dist ))+
  geom_point(alpha = .05)+
  geom_smooth(method="lm")

model <- lm(diff_from_sn_avg_ftpct ~ diff_from_sn_avg_dist, data = gbg_player_w_distance )
summary(model)

## Offense isolated

model <- lm(diff_from_sn_avg_fgpct ~ diff_from_sn_avg_off,
data = gbg_player_w_distance_at )
summary(model)

model <- lm(diff_from_sn_avg_fg3pct ~ diff_from_sn_avg_off,
data = gbg_player_w_distance_at3 )
summary(model)

model <- lm(diff_from_sn_avg_ftpct ~ diff_from_sn_avg_off, 
data = gbg_player_w_distance )
summary(model)

```

**Description:**  
Distance and the performance metrics are deviating from their season averages 
in tandem. To be clear this is not causal aka more distance did not necessarily 
make players shoot better nor does better shooting make them cover more 
distance in a game. All I can assert with the regressiions below is they are 
correlated.  

I only used percentage based summaries because total statistics like rebounds
are obviously linked to distance simply by playing longer.

**Whole game:**   

*difference_from_season_avg_fg* = $\beta_{0}$ + $\beta_{1}$ (*difference_from_season_avg_distance*)  
p-value $\beta_{1}$ : <2e-16 (t statistic 22.30 very strong)  

*difference_from_season_avg_fg3* = $\beta_{0}$ + $\beta_{1}$ (*difference_from_season_avg_distance*)  
p-value $\beta_{1}$ : <2e-16 (t statistic 11.53 very strong)   

*difference_from_season_avg_ft* = $\beta_{0}$ + $\beta_{1}$ (*difference_from_season_avg_distance*)  
p-value $\beta_{1}$ : 0.00512 (Not considering significant)   


**Offense Isolated:**  

*difference_from_season_avg_fg* = $\beta_{0}$ + $\beta_{1}$(*difference_from_season_avg_dist_off*)  
p-value $\beta_{1}$ : <2e-16 (t statistic 19.69 very strong)

*difference_from_season_avg_fg3* = $\beta_{0}$ +$\beta_{1}$(*difference_from_season_avg_dist_off*)  
p-value $\beta_{1}$ : <2e-16 (t statistic 10.714 very strong)

*difference_from_season_avg_ft* = $\beta_{0}$ + $\beta_{1}$(*difference_from_season_avg_dist_off*)  
p-value $\beta_{1}$ : 0013 (Not considering significant)


### Speed

An alternative hypothesis could be players who are "hot" get left in the game so 
they cover more distance. To check this speed will adjust for the players 
time in game. However, it may no longer act as a proxy for fatigue because it is 
agnostic to total exertion. Speed is a little like the derivative and total 
distance the integral over effort in the game.

```{r include = FALSE}
gbg_player_w_distance_at3 <-  gbg_player_w_distance %>%
  filter(fg3a > 0 )

gbg_player_w_distance_at <-  gbg_player_w_distance %>%
  filter(fga > 0 )

gbg_player_w_distance_at %>%
ggplot(aes(x=diff_from_sn_avg_fgpct, y =diff_from_sn_avg_speed ))+
  geom_point(alpha = .05)+
  geom_smooth(method="lm") 

model <- lm(diff_from_sn_avg_fgpct ~ diff_from_sn_avg_speed, 
            data = gbg_player_w_distance_at )
summary(model)


model <- lm(diff_from_sn_avg_fg3pct ~ diff_from_sn_avg_speed, 
            data = gbg_player_w_distance_at3 )
summary(model)

model <- lm(diff_from_sn_avg_ftpct ~ diff_from_sn_avg_speed, 
            data = gbg_player_w_distance )
summary(model)

model <- lm(diff_from_sn_avg_oreb ~ diff_from_sn_avg_speed, 
            data = gbg_player_w_distance )
summary(model)

model <- lm(diff_from_sn_avg_tov ~ diff_from_sn_avg_speed, 
            data = gbg_player_w_distance )
summary(model)

model <- lm(diff_from_sn_avg_stl ~ diff_from_sn_avg_speed, 
            data = gbg_player_w_distance )
summary(model)

```

The relationship flips at this level which is not consistent with the 
interpretation from above. All 4 significant relationships show playing faster
is correlate with worse outcomes.

**Whole game:**

*diff_from_sn_avg_fgpct* = $\beta_{0}$ - $\beta_{1}$(*diff_from_sn_avg_speed*)  
p-value  $\beta_{1}$: <2e-07 (t-value -5.144)  

*diff_from_sn_avg_fg3pct* =$\beta_{0}$ - $\beta_{1}$(*diff_from_sn_avg_speed*)  
p-value  $\beta_{1}$: 1.94e-06 (t-value -4.762 ) 

*diff_from_sn_avg_tov* =$\beta_{0}$ + $\beta_{1}$(*diff_from_sn_avg_speed*)  
p-value  $\beta_{1}$: 0.0037 (t-value 2.903) 

*diff_from_sn_avg_tov* =$\beta_{0}$ + $\beta_{1}$(*diff_from_sn_avg_speed*)  
p-value  $\beta_{1}$: 0.000353 (t-value 3.574 probably no relationship) 
 
*diff_from_sn_avg_oreb* No relationship p-vlaue > .1  
*diff_from_sn_avg_ft* No relationship p-vlaue > .1  

Alternative interpretation:
Perhaps there is too much randomness introduced when only filtering for playing 
15 seconds in the game. To check I filtered for played 15 minutes in the game
and this reduced the data from 24k observations to 19k.

```{r include = FALSE}
#### Repeated distance ############################################# 

gbg_player_w_distance <- gbg_player_w_distance %>% filter(min > 15)

##fgpct all distance
gbg_player_w_distance_at <-  gbg_player_w_distance %>%
  filter(fga > 0 )

gbg_player_w_distance %>%
ggplot(aes(x=diff_from_sn_avg_fgpct, y =diff_from_sn_avg_dist ))+
  geom_point(alpha = .05)+
  geom_smooth(method="lm")

model <- lm(diff_from_sn_avg_fgpct ~ diff_from_sn_avg_dist, 
            data = gbg_player_w_distance_at )
summary(model)

# fg3pct all distance
gbg_player_w_distance_at3 <-  gbg_player_w_distance %>%
  filter(fg3a > 0 )

gbg_player_w_distance_at3 %>%
ggplot(aes(x=diff_from_sn_avg_fg3pct, y =diff_from_sn_avg_dist ))+
  geom_point(alpha = .05)+
  geom_smooth(method="lm")

model <- lm(diff_from_sn_avg_fg3pct ~ diff_from_sn_avg_dist, 
            data = gbg_player_w_distance_at3 )
summary(model)

# ftpct all distance
gbg_player_w_distance %>%
  filter(!is.na(ftpct)) %>%
ggplot(aes(x=diff_from_sn_avg_ftpct, y =diff_from_sn_avg_dist ))+
  geom_point(alpha = .05)+
  geom_smooth(method="lm")

model <- lm(diff_from_sn_avg_ftpct ~ diff_from_sn_avg_dist, data = gbg_player_w_distance )
summary(model)

## Offense isolated

model <- lm(diff_from_sn_avg_fgpct ~ diff_from_sn_avg_off,
data = gbg_player_w_distance_at )
summary(model)

model <- lm(diff_from_sn_avg_fg3pct ~ diff_from_sn_avg_off,
data = gbg_player_w_distance_at3 )
summary(model)

model <- lm(diff_from_sn_avg_ftpct ~ diff_from_sn_avg_off, 
data = gbg_player_w_distance )
summary(model)


####  Repeat with speed  ##################################

gbg_player_w_distance_at3 <-  gbg_player_w_distance %>%
  filter(fg3a > 0 )

gbg_player_w_distance_at <-  gbg_player_w_distance %>%
  filter(fga > 0 )

gbg_player_w_distance_at %>%
ggplot(aes(x=diff_from_sn_avg_fgpct, y =diff_from_sn_avg_speed ))+
  geom_point(alpha = .05)+
  geom_smooth(method="lm") 

model <- lm(diff_from_sn_avg_fgpct ~ diff_from_sn_avg_speed, 
            data = gbg_player_w_distance_at )
summary(model)


model <- lm(diff_from_sn_avg_fg3pct ~ diff_from_sn_avg_speed, 
            data = gbg_player_w_distance_at3 )
summary(model)

model <- lm(diff_from_sn_avg_ftpct ~ diff_from_sn_avg_speed, 
            data = gbg_player_w_distance )
summary(model)

model <- lm(diff_from_sn_avg_oreb ~ diff_from_sn_avg_speed, 
            data = gbg_player_w_distance )
summary(model)

model <- lm(diff_from_sn_avg_tov ~ diff_from_sn_avg_speed, 
            data = gbg_player_w_distance )
summary(model)

model <- lm(diff_from_sn_avg_stl ~ diff_from_sn_avg_speed, 
            data = gbg_player_w_distance )
summary(model)


```

#### Same analysis as above repeated for speed and distance.
* Correlation between distance and fg percent and fg3 percent became slightly 
stronger. 
* All of the flipped relationships with speed now disapear. No idea if my 
hypothesis is correct

**Whole game:**

*diff_from_sn_avg_fgpct* = $\beta_{0}$ - $\beta_{1}$(*diff_from_sn_avg_speed*)   
p-value  $\beta_{1}$:  0.461 (t-value -0.737)  

*diff_from_sn_avg_fg3pct* =$\beta_{0}$ - $\beta_{1}$(*diff_from_sn_avg_speed*)   
p-value  $\beta_{1}$: 0.0119 (t-value -2.516 ) 

*diff_from_sn_avg_tov* =$\beta_{0}$ + $\beta_{1}$(*diff_from_sn_avg_speed*)   
p-value  $\beta_{1}$: 0.436 (t-value 0.779)  

*diff_from_sn_avg_stl* =$\beta_{0}$ + $\beta_{1}$(*diff_from_sn_avg_speed*)   
p-value  $\beta_{1}$: 6.26e-08 (t-value 5.413)  

My takeaway from this is speed does not offer any information. This is 
probably for the better as it resolves the issues with distance. It is also good
the distance relationships got stronger 

## 2.  
I think at this point most of the analysis we can conduct to validate distance 
as our proxy has been done.



```{r include = FALSE}
## investigating plusminus and distance but it is too noisy.
## put on hold and not included in markdown 
# 
# players_5_plus_min <- gbg_player_w_distance %>% filter(min >5)
# 
# players_5_10_min <- gbg_player_w_distance %>% filter(min >5 & min <10)
# 
# players_40_plus_min <- gbg_player_w_distance %>% filter(min >40)
# 
# players_5_plus_min %>% 
#   ggplot(aes(x=avg_speed,y =plusminus ))+
#   geom_point(alpha = .05)+
#   geom_smooth(method="lm")+
#   labs(title = "all players with 5+ min played")+
#   xlim(3,5.5)+
#   ylim(-30,30)
# 
# players_5_10_min %>% 
#   ggplot(aes(x = avg_speed,y = plusminus ))+
#   geom_point(alpha = .5)+
#   geom_smooth(method="lm")+
#   labs(title = "Players with 5-10 min in the game")+
#   xlim(2.75,5.5)+
#   ylim(-40,40)
# 
# players_40_plus_min %>% 
#   ggplot(aes(x=avg_speed,y =plusminus ))+
#   geom_point(alpha = .75)+
#   geom_smooth(method="lm")+
#   labs(title = "players with 40+ minutes in the game")+
#   xlim(2.75, 5.5)+
#   ylim(-40,40)
# 
# 
# ## model 
# model0 <- lm(plusminus ~ avg_speed, data = players_5_plus_min)
# #summary(model0)
# 
# model1 <- lm(plusminus ~ avg_speed, data = players_5_10_min)
# #summary(model1)
# 
# model2 <- lm(plusminus ~ avg_speed, data = players_40_plus_min)
# #summary(model2)


# The relationship of plusminus ~ avg_speed for all players with 5 or more minutes 
# in the game is: plusminus = 11.429 -2.33(avg_speed)  
# p-value: <2e-16

```

```{r fig.height =6, include = FALSE }
## A test space add include = FALSE before final knit


library("ggcorrplot") 

## offense correlation matrix
data_for_corr_offense <- gbg_player_w_distance %>% 
  dplyr::select( diff_from_sn_avg_speed_off, diff_from_sn_avg_off,
                 diff_from_sn_avg_fgpct, 
                 diff_from_sn_avg_fg3pct, diff_from_sn_avg_oreb,
                 diff_from_sn_avg_stl, diff_from_sn_avg_tov) %>%
  filter(!is.na(diff_from_sn_avg_speed_off) & !is.na(diff_from_sn_avg_off))

round_cor_matrix <- 
  round(cor(data_for_corr_offense), 2)

ggcorrplot(round_cor_matrix, 
           hc.order = TRUE,
           type = "lower",
           lab = TRUE)
## Exploring corr matrix stuff
# fg_pct shift 

## Defenese correlation matrix

## Saturated model 
sat_mod <- lm(plusminus ~ fgm + fga + pctFG + fg3m + fg3a + pctFG3 + pctFT + 
                fg2m + fg2a +pctFG2 +ftm + fta +oreb +dreb + treb + ast +stl +
                blk + tov+ pf, data =gbg_player_w_distance )
summary(sat_mod)

## removed attempts and made 
reduc_mod_1 <- lm(plusminus ~ pctFG + pctFG3 + pctFT + 
                 + pctFG2 + ftm + fta +oreb +dreb + treb + ast +stl +
                blk + tov+ pf, data =gbg_player_w_distance )





```












