---
title: "Follow up on Windows from Yesterday"
author: "Austin Stephen"
date: "7/21/2021"
output: html_document
---
```{r include = FALSE}
library(nbastatR)
library(NBAr)
library(ballr)
library(tidyverse)
library(dplyr)
knitr::opts_chunk$set(
  cache = TRUE,
  echo = FALSE,
  message = FALSE, 
  warning = FALSE,
  fig.height=3
)

## game by game player distance joined to player box scores.
gbg_player_w_distance <- read_csv(
  "../data/in_game_distance_data/in_game_dist_box_join.csv")

## windows of time analysis 
player_Window_dist_daily_2014_15 <- read_csv(
  "../data/in_game_distance_data/player_Window_dist_daily_2014_15.csv")%>%
  select("player_id","date","dist_5dayWindow","dist_3dayWindow","dist_7dayWindow",
         "dist_10dayWindow")

layer_Window_speed_daily_2014_15 <- read_csv(
  "../data/in_game_distance_data/player_Window_speed_daily_2014_15.csv") %>%
  select("player_id","date","speed_5dayWindow","speed_7dayWindow")

player_Window_dist_daily_2014_15$date <- as.Date(
  player_Window_dist_daily_2014_15$date, "%m/%d/%Y")

layer_Window_speed_daily_2014_15$date <- as.Date(
  layer_Window_speed_daily_2014_15$date, "%m/%d/%Y")


## removing all instances where player played less than 15 seconds because 
## these are just not insightful instances 
gbg_player_w_distance <- gbg_player_w_distance %>% filter(min > .25)


## window
gbg_player_w_distance <- merge(x = gbg_player_w_distance,
                               y = player_Window_dist_daily_2014_15,
                               by.x =  c("player_id","date" ),
                               by.y = c("player_id","date" ))

gbg_player_w_distance <- merge(x = gbg_player_w_distance,
                               y = layer_Window_speed_daily_2014_15,
                               by.x =  c("player_id","date" ),
                               by.y = c("player_id","date" ))

## Adding a difference from mean 5 day window
gbg_player_w_distance <- gbg_player_w_distance %>%
  group_by(player_id) %>%
  mutate(sn_avg_wind_3_dist = mean(dist_3dayWindow)) %>%
  mutate(diff_sn_avg_wind_3_dist = dist_3dayWindow - sn_avg_wind_3_dist ) %>%
  mutate(sn_avg_wind_5_dist = mean(dist_5dayWindow)) %>%
  mutate(diff_sn_avg_wind_5_dist = dist_5dayWindow - sn_avg_wind_5_dist ) %>%
  mutate(sn_avg_wind_7_dist = mean(dist_7dayWindow)) %>%
  mutate(diff_sn_avg_wind_7_dist = dist_7dayWindow - sn_avg_wind_7_dist ) %>%
  mutate(sn_avg_wind_10_dist = mean(dist_10dayWindow)) %>%
  mutate(diff_sn_avg_wind_10_dist = dist_10dayWindow - sn_avg_wind_10_dist ) %>%
  mutate(sn_avg_wind_5_speed = mean(speed_5dayWindow)) %>%
  mutate(diff_sn_avg_wind_5_speed = speed_5dayWindow - sn_avg_wind_5_speed ) %>%
  mutate(sn_avg_wind_7_speed = mean(speed_7dayWindow)) %>%
  mutate(diff_sn_avg_wind_7_speed = speed_7dayWindow - sn_avg_wind_7_speed )



gbg_player_w_distance_play_time <- gbg_player_w_distance %>% filter(min > 10)

## Between game
between_game_dist <- read_csv("../data/between_game_dist/between_game_dist.csv") %>%
  rename(distance_team_traveled_pre_game = "Distance") %>%
  select(-c("Season","Date","Team", "Opponent","Month","Week"))

gbg_player_w_distance_subset <- merge(x = gbg_player_w_distance,
                               y = between_game_dist,
                               by.x = c("idGame","team_id"),
                               by.y = c("game_id","team_id")
                               )




```
### TLDR:
1. Running win percent difference was an unfair predictor because it was
inclusive  of the observation so it incorporated the response into the running 
total. I corrected this in the comparison below.  
2. We have a pretty strong proxy for team strength in my eyes. 
3. TODO: Need to build a model over the residuals still but have done analysis by adding 
fatigue factors as additional predictors (Matthew has this)   
4. TODO: Random forest and gradient boosted trees over an aggregation of travel 
data and PDP analysis of importance.  
5. QUESTION: Worth our time to check 12 and 14 game window sizes centered since 
it appears to get better?  

### Speed mentioned in last windows rmd
* This is every player in every game of the 2014-15 season  
* Very clear any relationship with in-game distance from my window rmd was just due
to play time  
```{r fig.height = 2, fig.width= 3}

gbg_player_w_distance %>%
ggplot(aes(x = diff_sn_avg_wind_5_speed,y = diff_from_sn_avg_speed))+
  geom_point(alpha = .05)+
  geom_smooth(method="lm")+ 
  labs(y = "diff from season avg dist per game",
       x = "diff from season avg 3 day window")

gbg_player_w_distance %>%
ggplot(aes(x = diff_sn_avg_wind_7_speed,y = diff_from_sn_avg_speed))+
  geom_point(alpha = .05)+
  geom_smooth(method="lm")+ 
  labs(y = "diff from season avg dist per game",
       x = "diff from season avg 3 day window")+
  scale_x_continuous(limits = c(-11,11))+
  scale_y_continuous(limits = c(-2.5,2.5))

```


### Accounting for team strength
* This is a pretty good result I think
```{r fig.height= 3}
together_net_Rating_allWindows <- read_csv("../data/together_net_Rating_allWindows_15_30.csv") %>%
   # select(c("Team","Opponent", "Date", "game_id","season", "net_rating_diff","opp_net_rating",
   #          "w_lpercent", "game_net_rating", "net_rating_2gameWindow", "net_rating_5gameWindow",
   #          "net_rating_7gameWindow", "net_rating_10gameWindow", "opp_net_rating_2gameWindow",
   #          "opp_net_rating_5gameWindow","opp_net_rating_7gameWindow","opp_net_rating_10gameWindow",
   #         "opp_net_rating_CentergameWindow", "net_rating_CentergameWindow")) %>%
  mutate(wind_2 = net_rating_2gameWindow - opp_net_rating_2gameWindow,
         wind_5 = net_rating_5gameWindow - opp_net_rating_5gameWindow,
         wind_7 = net_rating_7gameWindow - opp_net_rating_7gameWindow,
         wind_10 = net_rating_10gameWindow - opp_net_rating_10gameWindow,
         wind_15 = net_rating_15gameWindow - opp_net_rating_15gameWindow,
         wind_20 = net_rating_20gameWindow - opp_net_rating_20gameWindow,
         wind_25 = net_rating_25gameWindow - opp_net_rating_25gameWindow,
         wind_30 = net_rating_30gameWindow - opp_net_rating_30gameWindow,
         wind_10_center = net_rating_CentergameWindow - opp_net_rating_CentergameWindow )

together_net_Rating_allWindows <- together_net_Rating_allWindows %>%
  mutate( hollow_wind5_home = net_rating_5gameWindow - net_rating_2gameWindow,
          hollow_wind5_opp = opp_net_rating_5gameWindow - opp_net_rating_2gameWindow,
          wind_5_hollow = hollow_wind5_home - hollow_wind5_opp,
          hollow_wind7_home = net_rating_7gameWindow - net_rating_5gameWindow,
          hollow_wind7_opp = opp_net_rating_7gameWindow - opp_net_rating_5gameWindow,
          wind_7_hollow = hollow_wind7_home - hollow_wind7_opp,
          hollow_wind10_home = net_rating_10gameWindow - net_rating_7gameWindow,
          hollow_wind10_opp = opp_net_rating_10gameWindow - opp_net_rating_7gameWindow,
          wind_10_hollow = hollow_wind10_home - hollow_wind10_opp,
          hollow_wind15_home = net_rating_15gameWindow - net_rating_10gameWindow,
          hollow_wind15_opp = opp_net_rating_15gameWindow - opp_net_rating_10gameWindow,
          wind_15_hollow = hollow_wind15_home - hollow_wind15_opp,
          hollow_wind20_home = net_rating_20gameWindow - net_rating_15gameWindow,
          hollow_wind20_opp = opp_net_rating_20gameWindow - opp_net_rating_15gameWindow,
          wind_20_hollow = hollow_wind20_home - hollow_wind20_opp,
          hollow_wind25_home = net_rating_25gameWindow - net_rating_20gameWindow,
          hollow_wind25_opp = opp_net_rating_25gameWindow - opp_net_rating_20gameWindow,
          wind_25_hollow = hollow_wind25_home - hollow_wind25_opp,
          hollow_wind30_home = net_rating_30gameWindow - net_rating_25gameWindow,
          hollow_wind30_opp = opp_net_rating_30gameWindow - opp_net_rating_25gameWindow,
          wind_30_hollow = hollow_wind30_home - hollow_wind30_opp,
          hollow_wind10_home_5 = net_rating_10gameWindow - net_rating_5gameWindow,
          hollow_wind10_opp_5 = opp_net_rating_10gameWindow - opp_net_rating_5gameWindow,
          wind_10_hollow_5 = hollow_wind10_home_5 - hollow_wind10_opp_5
          )
   # select(c("Team","Opponent", "Date", "game_id","season", "net_rating_diff","opp_net_rating",
   #          "w_lpercent", "game_net_rating", "net_rating_2gameWindow", "net_rating_5gameWindow",
   #          "net_rating_7gameWindow", "net_rating_10gameWindow", "opp_net_rating_2gameWindow",
   #          "opp_net_rating_5gameWindow","opp_net_rating_7gameWindow","opp_net_rating_10gameWindow",
   #         "opp_net_rating_CentergameWindow", "net_rating_CentergameWindow", 
   #         "wind_30_hollow", "wind_25_hollow" ))
  

together_net_Rating_allWindows <- together_net_Rating_allWindows %>%
  group_by("Team","season") %>%
  mutate(win_percent_diff = ifelse(score_diff >0,w.x-1,w.x) /pmax(1,w.x + l.x -1))
```

```{r fig.height= 3}
## Running win percent diff
# summary(lm(score_diff ~ win_percent_diff, data = together_net_Rating_allWindows))
# summary(lm(score_diff ~ win_percent_diff + net_rating_diff , data = together_net_Rating_allWindows))
# #Adjusted R-squared:  0.05231 => 0.1099 
# 
# ## Window size of 2
# summary(lm(score_diff ~ wind_2, data = together_net_Rating_allWindows))
# summary(lm(score_diff ~ wind_2 + net_rating_diff , data = together_net_Rating_allWindows))
# #Adjusted R-squared:  0.04534  =>  0.1106
# 
# 
# ## Window size of 5
# summary(lm(score_diff ~ wind_5, data = together_net_Rating_allWindows))
# summary(lm(score_diff ~ wind_5 + net_rating_diff , data = together_net_Rating_allWindows))
# #Adjusted R-squared:  0.084 => 0.1272
# 
# ## Window size of 7
# summary(lm(score_diff ~ wind_7, data = together_net_Rating_allWindows))
# summary(lm(score_diff ~ wind_7 + net_rating_diff , data = together_net_Rating_allWindows))
# # Adjusted R-squared:  0.1007 => 0.1357
# 
# ## Window size of 10
# summary(lm(score_diff ~ wind_10, data = together_net_Rating_allWindows))
# summary(lm(score_diff ~ wind_10 + net_rating_diff , data = together_net_Rating_allWindows))
# # Adjusted R-squared: 0.1143 => 0.1426
# 
# ## Window size of 15
# summary(lm(score_diff ~ wind_15, data = together_net_Rating_allWindows))
# summary(lm(score_diff ~ wind_15 + net_rating_diff , data = together_net_Rating_allWindows))
# # Adjusted R-squared: 0.1267  => 0.1486 
# 
# ## Window size of 20
# summary(lm(score_diff ~ wind_20, data = together_net_Rating_allWindows))
# summary(lm(score_diff ~ wind_20 + net_rating_diff , data = together_net_Rating_allWindows))
# # Adjusted R-squared: 0.1314  => 0.1501 
# 
# ## Window size of 25
# summary(lm(score_diff ~ wind_25, data = together_net_Rating_allWindows))
# summary(lm(score_diff ~ wind_25 + net_rating_diff , data = together_net_Rating_allWindows))
# # Adjusted R-squared: 0.1336  => 0.1506
# 
# ## Window size of 30
# summary(lm(score_diff ~ wind_30, data = together_net_Rating_allWindows))
# summary(lm(score_diff ~ wind_30 + net_rating_diff , data = together_net_Rating_allWindows))
# # Adjusted R-squared: 0.1351  =>  0.1511

# ## Window size 10 hollow
# summary(lm(score_diff ~ wind_2 + wind_5_hollow + wind_7_hollow + wind_10_hollow, 
#            data = together_net_Rating_allWindows))
# summary(lm(score_diff ~ wind_2 + wind_5_hollow + wind_7_hollow + wind_10_hollow + net_rating_diff , 
#            data = together_net_Rating_allWindows))
# Adjusted R-squared: 0.1145 => 0.1427

## Window size of 30 hollow
# summary(lm(score_diff ~ wind_5 + wind_10_hollow_5 + wind_15_hollow + wind_20_hollow +
#              wind_25_hollow + wind_30_hollow, data = together_net_Rating_allWindows))
# summary(lm(score_diff ~ wind_5 + wind_10_hollow_5 + wind_15_hollow + wind_20_hollow +
#              wind_25_hollow + wind_30_hollow + net_rating_diff , data = together_net_Rating_allWindows))
# Adjusted R-squared:  0.140  =>  0.1552 

# ## Window size of 10 centered
# summary(lm(score_diff ~ wind_10_center, data = together_net_Rating_allWindows))
# summary(lm(score_diff ~ wind_10_center + net_rating_diff , data = together_net_Rating_allWindows))
# # Adjusted R-squared: 0.1241 => 0.1487
# 


## Not a big enough payout for the increase in complexity- important for the 
# paper because it shows a system that used a contious weighting approachwould not work
r_squared_window <- c(0.05231, 0.04534, 0.0840, 0.1007, 0.1143, 0.1267, 0.1314,
                      0.1336, 0.1351, 0.1145,0.140, 0.1241)

r_squared_window_plus_sn_avg <- c(0.1099, 0.1106, 0.1272, 0.1357, 0.1426, 
                            0.1486, 0.1501, 0.1506, 0.1511, 0.1427,0.1552, 0.1487)
diff <- r_squared_window_plus_sn_avg - r_squared_window

window_size <- c("win_pct" , "2" , "5" , "7" ,"10" ,"15", "20", "25", "30", 
                 "10*","30*", "10_cen")
window_size <- ordered(window_size, levels = c("win_pct" , "2" , "5" , "7" ,
                        "10" ,"15", "20", "25", "30", "10*","30*", "10_cen"))

plot_data <- data.frame(r_squared_window,r_squared_window_plus_sn_avg, window_size, diff)

plot_data %>% ggplot()+
  geom_point(y=r_squared_window, x=window_size, aes(color = "window"))+
  geom_point(y=r_squared_window_plus_sn_avg, x= window_size, aes(color = "window & season_avg"))+
  geom_point(y=diff, x= window_size, aes(color = "additional varriance"))+
  geom_line(aes(y=r_squared_window,x=window_size ), group = 1, color = "green")+
  geom_line(aes(y=r_squared_window_plus_sn_avg, x= window_size), group = 1, color = "darkturquoise")+
  geom_line(aes(y=diff, x= window_size), group = 1, color = "red")+
  labs(color = "Predictor",
       y= "r-squared",
       x ="Window size",
       caption= "* each subset of prior games got its own coeff",
       title = "Prefromance on Entire Dataset")+
  theme_bw()



```

More details on 10*:  I made the past 10 games window into a partition
of the space using Grace's buckets (2, 5, 7, 10) taking out duplication.
Then I gave each new bucket its own coefficient with the intent of incorporating
temporal information about the games into the prediction. To be more concrete,
1 and 2 games back are a variable, 3,4 and 5 games back their own variable, 6 and 7 games back,
and 8-10 are their own variable in a linear model. This allows the regression to assign
coefficients that give decreasing importance to buckets of games as they become
more distant in time. It is essentially using "buckets" to allow the regression
to find how game importance diminishes over time as opposed to me defining some
heuristic. I personally don't think the complexity of this approach is worth the
payout in r-squared. However, I also think buckets will probably generalize
better out of sample than the continuous version of this idea from the last meeting.
Also interesting, is it accounts for about the same variance as the centered
window but it does not take from the overall average in accounting for this variance.

