---
title: "Why is less distance better?"
author: "Austin Stephen"
date: "7/8/2021"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE }
library(tidyverse)
library(dplyr)
library(nbastatR)
library(NBAr)
library(ggrepel)
knitr::opts_chunk$set(
  cache = TRUE,
  echo = FALSE,
  message = FALSE, 
  warning = FALSE,
  fig.height=3
)
## importing data 

## one test season 
game_by_game_distance_2019_season <-read.csv(
  "../data/in_game_distance_data/tracking_by_player_season_daily_2019_20.csv")

## Player id and position map table 
player_table <- read.csv("../data/player_tables/player_table_all_2010_to_21.csv")
  
## player tracking distance 
player_tracking <- read.csv(
  "../data/in_game_distance_data/tracking_by_player_season_summaries_2013_to_20.csv")

## team tracking distance 
team_tracking <- read.csv(
  "../data/in_game_distance_data/tracking_by_team_season_summaries_2013_to_20.csv")

## Player season box scores
season_box_scores_player_2010_20 <- read.csv(
  "../data/box_scores/season_box_scores_player_2010_20.csv")

## Team season box scores
season_box_scores_team_2010_20 <- read.csv(
  "../data/box_scores/season_box_scores_team_2010_20.csv")

season_box_scores_team_2010_20$season

## Joining team box scores to season distance data 
team_tracking_w_box_scsores <- merge(team_tracking,
                                    season_box_scores_team_2010_20,
                                    by.x = c("team_id","season"), 
                                    by.y = c("team_id","season")) %>%
  select(-c("team_name.y","gp.y","w.y","l.y", "min1")) %>%
  rename( "w" = w.x, "gp" = "gp.x", "l" = l.x)

```
## TLDR:
1. Offense distance accounts for most of the negative relationship with wins.
2. Three point percentage is not the cause.
3. Field goal percentage is not the cause.
4. Offensive rebounds are not the cause.
5. Turnovers are not the cause.  
6. __(Most Interesting)__ There is a strong positive correlation with 
assists per game and wins. 
Also, there is a strong positive correlation with assists per game and distance. 
This is the opposite of what we observed with wins and distance per game.
Meaning this had to be outweighed by some other factor I can't find for the 
negative relationship of interest to exist.

## Why is there an inverse correlation with wins and distance?  
```{r }
## Teams are covering more distance per game but players are not postulate this 
#  relationship with juxtaposing those visuals against each other and means. 

ggplot(team_tracking, aes(x = dist_miles, y= w)) +
    geom_point(size=.5, alpha = .75)+
    theme_bw()+
  facet_wrap(~season) +
  geom_smooth(method = "lm", se = FALSE)+
    labs(title = "Teams wins and average per distance per game",
         x = " distance (miles)")


## linear model distance and wins
model <- lm(w ~ dist_miles, data = team_tracking)

team_tracking %>% 
  ggplot(aes(x=dist_miles, y= w)) + 
  geom_point()+
  geom_abline(intercept = 121.6, slope = -4.73, color = "darkred")+
  theme_bw()+ 
  labs( title = "2013 to 2020 wins and distance")

```

y = 121.6 + -4.73  
p-value: .00018 and adj r-squared: .053  

### 2018  
```{r}
## making subset of teams to look at
team_tracking_2018 <- team_tracking %>% 
  filter(season == 2018)

team_tracking_2018_subset <- team_tracking_2018 %>% 
  filter(team_abbreviation %in% c( "NYK", "MIL", "TOR","HOU","SAC"))

player_tracking_2018_subset <- player_tracking %>% 
  filter(season == 2018) %>% 
  filter(team_abbreviation %in% c( "NYK", "MIL", "TOR","HOU","SAC"))

## full plot
ggplot(data =team_tracking_2018, aes(x = dist_miles, y= w)) +
    geom_point(data = team_tracking_2018, size = 2,color = "black")+
    geom_point(data = team_tracking_2018_subset, size = 3, color = "darkred")+
    coord_cartesian(xlim = c(16, 19), ylim = c(0, 82)) +
    scale_color_brewer(palette="Set1")+
    geom_text(aes(
      label=ifelse(team_abbreviation %in% c( "NYK", "MIL", "TOR","HOU","SAC"),
                   team_abbreviation,'')),
      hjust=0, vjust=0)+
    theme_bw()


```

```{r}
## distribution of players on teams
player_tracking_2018_subset %>% 
  ggplot(aes(y=dist_miles, x= team_abbreviation))+
  geom_violin()+
  geom_point()+
  theme_bw()

player_tracking_2018_subset %>% 
  filter(team_abbreviation %in% c("HOU")) %>%
  ggplot(aes(y=dist_miles, x= team_abbreviation))+
  geom_point()+
  geom_text_repel(aes(label = player_name),
                  box.padding = 0.5,
                  point.padding = .01,
                  label.size = 0)+
  
  theme_bw()
```

HOU the best team by record has a cluster of players that run less than .5 a miles a game.
However, NYK has no players who run under .75 miles a game.

### 2017 

```{r}
## making subset of teams to look at
team_tracking_2017 <- team_tracking %>% 
  filter(season == 2017)

team_tracking_2017_subset <- team_tracking_2017 %>% 
  filter(team_abbreviation %in% c( "HOU", "TOR", "CHI","PHI","MIN"))

player_tracking_2017_subset <- player_tracking %>% 
  filter(season == 2017) %>% 
  filter(team_abbreviation %in% c( "HOU", "TOR", "CHI","PHI","MIN"))

## full plot
ggplot(data =team_tracking_2017, aes(x = dist_miles, y= w)) +
    geom_point(data = team_tracking_2017, size = 2,color = "black")+
    geom_point(data = team_tracking_2017_subset, size = 3, color = "darkred")+
    coord_cartesian(xlim = c(16, 19), ylim = c(0, 82)) +
    scale_color_brewer(palette="Set1")+
    geom_text(aes(
      label=ifelse(team_abbreviation %in% c( "HOU", "TOR", "CHI","PHI","MIN"),
                   team_abbreviation,'')),
       hjust=0, vjust=0)+
    theme_bw()

## distribution of players on teams
player_tracking_2017_subset %>% 
  ggplot(aes(y=dist_miles, x= team_abbreviation))+
  geom_violin()+
  geom_point()+
  theme_bw()

player_tracking_2017_subset %>% 
  filter(team_abbreviation %in% c("MIN")) %>%
  ggplot(aes(y=dist_miles, x= team_abbreviation))+
  geom_point()+
  geom_text_repel(aes(label = player_name),
                  box.padding = 0.5,
                  point.padding = .01,
                  label.size = 0)+
  
  theme_bw()
```

I don't really see any relationship in 2017.

### 2016 

```{r}
## making subset of teams to look at
team_tracking_2016 <- team_tracking %>% 
  filter(season == 2016)

team_tracking_2016_subset <- team_tracking_2016 %>% 
  filter(team_abbreviation %in% c( "PHI", "LAC", "GSW","PHX","BKN"))

player_tracking_2016_subset <- player_tracking %>% 
  filter(season == 2016) %>% 
  filter(team_abbreviation %in% c( "PHI", "LAC", "GSW","PHX","BKN"))

## full plot
ggplot(data =team_tracking_2016, aes(x = dist_miles, y= w)) +
    geom_point(data = team_tracking_2016, size = 2,color = "black")+
    geom_point(data = team_tracking_2016_subset, size = 3, color = "darkred")+
    coord_cartesian(xlim = c(16, 19), ylim = c(0, 82)) +
    scale_color_brewer(palette="Set1")+
    geom_text(aes(
      label=ifelse(team_abbreviation %in% c( "PHI", "LAC", "GSW","PHX","BKN"),
                   team_abbreviation,'')),
       hjust=0, vjust=0)+
    theme_bw()

## distribution of players on teams
player_tracking_2016_subset %>% 
  ggplot(aes(y=dist_miles, x= team_abbreviation))+
  geom_violin()+
  geom_point()+
  theme_bw()

player_tracking_2016_subset %>% 
  filter(team_abbreviation %in% c("LAC")) %>%
  ggplot(aes(y=dist_miles, x= team_abbreviation))+
  geom_point()+
  geom_text_repel(aes(label = player_name),
                  box.padding = 0.5,
                  point.padding = .01,
                  label.size = 0)+
  
  theme_bw()
```

GSW was the best team and has no players under .5 so essentially the opposite of 
the 2019 trend with HOU.


### Offensive distance accounts for most of the association with wins
```{r}
ggplot(team_tracking, aes(x = dist_miles_off, y= w)) +
    geom_point(size=.5, alpha = .75)+
    theme_bw()+
  facet_wrap(~season) +
  geom_smooth(method = "lm", se = FALSE)+
    labs(title = "Offsense team's wins and average distance per game",
         x = " distance (mi)")
```

```{r}
ggplot(team_tracking, aes(x = dist_miles_def, y= w)) +
    geom_point(size=.5, alpha = .75)+
    theme_bw()+
  facet_wrap(~season) +
  geom_smooth(method = "lm", se = FALSE)+
    labs(title = "Defense Team's wins and average distance per game",
         x = " speed (mph)")
```

```{r}
## models offense and defense 

model <- lm(w ~ dist_miles_off, data = team_tracking)

team_tracking %>% 
  ggplot(aes(x=dist_miles_off, y= w)) + 
  geom_point()+
  geom_abline(intercept = 110.614, slope = -7.617, color = "darkred")+
  theme_bw()+ 
  labs( title = "2013 to 2020 wins and distance offense")
```

Linear model w ~ offense_distance  
p-value: 2.4 *10-5  
adj r-squared: 0.065  

```{r}
model <- lm(w ~ dist_miles_def, data = team_tracking)

team_tracking %>% 
  ggplot(aes(x=dist_miles_def, y= w)) + 
  geom_point()+
  geom_abline(intercept = 83.821, slope = -5.505, color = "darkred")+
  theme_bw()+ 
  labs( title = "2013 to 2020 wins and distance defense")
```

Linear model w ~ defense_distance  
p-value: 0.048   
adj r-squared: 0.012  

Significant relationship between distance on offense and wins. A negative 
or neutral relationship between distance on defense and wins. This means 
it is probably linked to some type of offensive statistic.  


## How distance relates with other offensive statistics  

### Three Point Percentage  

```{r}
team_tracking_w_box_scsores %>% 
  ggplot(aes(x=dist_miles_off, y= fg3_pct)) + 
  geom_point()+
  geom_abline(intercept = 110.614, slope = -7.617, color = "darkred")+
  theme_bw()+ 
  geom_smooth(method = "lm", se = FALSE) +
  labs( title = "2013 to 2020 three point shooting percentage and offensive distance")
```

```{r}
team_tracking_w_box_scsores %>% 
  ggplot(aes(x=fg3_pct, y= w)) + 
  geom_point()+
  theme_bw()+ 
  geom_smooth(method = "lm", se = FALSE) +
  labs( title = "2013 to 2020 three point shooting percentage and winning games")
```

There appears to be a slight positive or no relationship with three point shooting and 
distance but this is in the opposite direction of the trend with distance 
and wins. This is impressive because three point shooting is very strongly linked 
to wins which means the association we saw with distance overcame this relationship.

### Field Goal Percentage   

```{r}
team_tracking_w_box_scsores %>% 
  ggplot(aes(x=dist_miles_off, y= fg_pct)) + 
  geom_point()+
  geom_abline(intercept = 110.614, slope = -7.617, color = "darkred")+
  geom_smooth(method = "lm", se = FALSE) +
  theme_bw()+ 
  labs( title = "2013 to 2020 field goal percentage and offensive distance")

```

```{r}
team_tracking_w_box_scsores %>% 
  ggplot(aes(x=fg_pct, y= w)) + 
  geom_point()+
  geom_abline(intercept = 110.614, slope = -7.617, color = "darkred")+
  theme_bw()+ 
  geom_smooth(method = "lm", se = FALSE) +
  labs( title = "2013 to 2020 field goal percentage and wins")
```


Shooting percentage appears to have no relationship with team distance and is
clearly linked to winning games as expected.

### Offensive Rebounds  

```{r}
team_tracking_w_box_scsores %>% 
  ggplot(aes(x=dist_miles_off, y= oreb)) + 
  geom_point()+
  geom_abline(intercept = 110.614, slope = -7.617, color = "darkred")+
  geom_smooth(method = "lm", se = FALSE) +
  theme_bw()+ 
  labs( title = "2013 to 2020 offensive rebounds and offensive distance")
```

```{r}
team_tracking_w_box_scsores %>% 
  ggplot(aes(x=oreb, y= w)) + 
  geom_point()+
  geom_smooth(method = "lm", se = FALSE) +
  theme_bw()+ 
  labs( title = "2013 to 2020 offensive rebounds and wins")
```


Appears to be a slight negative correlation between offensive rebounds and 
distance which is initially promising. However, offensive rebounds has a negative 
association with wins so this acts to counteract the net effect of distance we are 
looking to explain. Also, we know getting offensive rebounds does not make 
a team worse so it must have to do with missing more shots enables more 
offensive rebounds. Overall, this does not help explain the effect.

### Turnovers

```{r}
team_tracking_w_box_scsores %>% 
  ggplot(aes(x=dist_miles_off, y= tov)) + 
  geom_point()+
  geom_smooth(method = "lm", se = FALSE) +
  theme_bw()+ 
  labs( title = "2013 to 2020 assists per game and offensive distance")

```

No relationship.


### Assists

```{r}
team_tracking_w_box_scsores %>% 
  ggplot(aes(x=dist_miles_off, y= ast)) + 
  geom_point()+
  geom_smooth(method = "lm", se = FALSE) +
  theme_bw()+ 
  labs( title = "2013 to 2020 assists per game and offensive distance")

model <- lm(ast ~ dist_miles_off, data = team_tracking_w_box_scsores)

```

```{r}
team_tracking_w_box_scsores %>% 
  ggplot(aes(x=ast, y= w)) + 
  geom_point()+
  geom_smooth(method = "lm", se = FALSE) +
  theme_bw()+ 
  labs( title = "2013 to 2020 assists per game and wins")

model <- lm(w ~ ast, data = team_tracking_w_box_scsores)
```


Assists and distance:  
p-value: 4.7 * 10^-8  
r-squared: 0.118  

Assists and wins:  
p-value: 5.854e-05
r-squared: 0.06574  

There is a significant positive relationship between distance and assists. 
Also, there is a significant positive relationship between assists and wins.
This is the opposite of what we would expect to see because it means distance 
had to overcome this relationship to get the negative correlation we saw.


### Total passes
```{r}
team_tracking_w_box_scsores %>% 
  ggplot(aes(x=dist_miles_off, y= ast)) + 
  geom_point()+
  geom_smooth(method = "lm", se = FALSE) +
  theme_bw()+ 
  labs( title = "2013 to 2020 assists per game and offensive distance")

model <- lm(ast ~ dist_miles_off, data = team_tracking_w_box_scsores)

```


