---
title: "Window analysis game-by-game and player level"
author: "Austin Stephen"
date: "7/19/2021"
output: html_document
---

```{r setup, include=FALSE }
library(nbastatR)
library(NBAr)
library(ballr)
library(tidyverse)
library(dplyr)
knitr::opts_chunk$set(
  cache = TRUE,
  echo = FALSE,
  message = FALSE, 
  warning = FALSE,
  fig.height=3
)

## game by game player distance joined to player box scores.
gbg_player_w_distance <- read_csv(
  "../data/in_game_distance_data/in_game_dist_box_join.csv")

## windows of time analysis 
player_Window_dist_daily_2014_15 <- read_csv(
  "../data/in_game_distance_data/player_Window_dist_daily_2014_15.csv")%>%
  select("player_id","date","dist_5dayWindow","dist_3dayWindow","dist_7dayWindow",
         "dist_10dayWindow")

player_Window_dist_daily_2014_15$date <- as.Date(
  player_Window_dist_daily_2014_15$date, "%m/%d/%Y")

## removing all instances where player played less than 15 seconds because 
## these are just not insightful instances 
gbg_player_w_distance <- gbg_player_w_distance %>% filter(min > .25)


## window
gbg_player_w_distance <- merge(x = gbg_player_w_distance,
                               y = player_Window_dist_daily_2014_15,
                               by.x =  c("player_id","date" ),
                               by.y = c("player_id","date" ))

## Adding a difference from mean 5 day window
gbg_player_w_distance <- gbg_player_w_distance %>%
  group_by(player_id) %>%
  mutate(sn_avg_wind_3_dist = mean(dist_3dayWindow)) %>%
  mutate(diff_sn_avg_wind_3_dist = dist_3dayWindow - sn_avg_wind_3_dist ) %>%
  mutate(sn_avg_wind_5_dist = mean(dist_5dayWindow)) %>%
  mutate(diff_sn_avg_wind_5_dist = dist_5dayWindow - sn_avg_wind_5_dist ) %>%
  mutate(sn_avg_wind_7_dist = mean(dist_7dayWindow)) %>%
  mutate(diff_sn_avg_wind_7_dist = dist_7dayWindow - sn_avg_wind_7_dist ) %>%
  mutate(sn_avg_wind_10_dist = mean(dist_10dayWindow)) %>%
  mutate(diff_sn_avg_wind_10_dist = dist_10dayWindow - sn_avg_wind_10_dist )



gbg_player_w_distance_play_time <- gbg_player_w_distance %>% filter(min > 10)

## Between game
between_game_dist <- read_csv("../data/between_game_dist/between_game_dist.csv") %>%
  rename(distance_team_traveled_pre_game = "Distance") %>%
  select(-c("Season","Date","Team", "Opponent","Month","Week"))

gbg_player_w_distance_subset <- merge(x = gbg_player_w_distance,
                               y = between_game_dist,
                               by.x = c("idGame","team_id"),
                               by.y = c("game_id","team_id")
                               )



## Distance differential at a given game

## Cross with Matthew's distance data to see if there is a correlation between the distance 
#  outside of a game and distance in game.

## See if teams adapt their styles to their opponents. So if defense distance is 
# correlated with opponent offense by season average and game.

## see if speed or distance differential is related to score score differential 

## Scatter plot by team that includes player name showing if having 1 outlier player 
#  leads to a decrease in the overall distance traveled.

```

### Take Aways:
* This is not capturing the refractory period due to fatigue (shows opposite actually). 
* This looks good for our rolling avg as a proxy for team strength as
current performance is closely tied to performance over the last 5 games. 
* Need to do new windows of speed because it is time adjusted and this might help
avoid players getting used more.

#### About the data:
Each observation is a single player in a single game. All players who were in 
the game for less than 15 seconds were removed. In total, It has 24k
observations from the 2014-15 regular season.

```{r fig.height=2, fig.width=3.5}
# No normalizing
gbg_player_w_distance %>%
ggplot(aes(x = dist_5dayWindow, y = dist_miles))+
  geom_point(alpha = .05)+
  geom_smooth(method="lm")+ 
  labs(y = "in-game distance (mi)",
       x = "Sum in-game dist 5 days prior to observation")

#summary(lm(dist_miles ~ dist_5dayWindow, data = gbg_player_w_distance))
# Adjusted for player averages
gbg_player_w_distance %>%
ggplot(aes(x = dist_5dayWindow,y = diff_from_sn_avg_dist))+
  geom_point(alpha = .05)+
  geom_smooth(method="lm")+ 
  labs(y = "diff from season avg dist",
       x = "Sum in-game dist 5 days prior to observation")

#summary(lm(diff_from_sn_avg_dist ~ dist_5dayWindow, data = gbg_player_w_distance))

# Adjusted for player averages and window averages 
gbg_player_w_distance %>%
ggplot(aes(x = diff_sn_avg_wind_5_dist,y = diff_from_sn_avg_dist))+
  geom_point(alpha = .05)+
  geom_smooth(method="lm")+ 
  labs(y = "diff from season avg dist",
       x = "diff from season avg 5 day window")

#summary(lm(diff_sn_avg_wind_5_dist ~ diff_from_sn_avg_dist, data = gbg_player_w_distance))

```

1. *dist_miles* = $\beta_{0}$ + $\beta_{1}$ (*Sum_of_dist_5_days_prior*)  
p-value $\beta_{1}$ : <2e-16 (t statistic 115.2)  

2. *diff_from_season_avg_distance* = $\beta_{0}$ + $\beta_{1}$
(*Sum_of_dist_5_days_prior*)  
p-value $\beta_{1}$ : <2e-16 (t statistic 31.93)  

3. *diff_from_season_avg_distance* = $\beta_{0}$ + $\beta_{1}$
(*diff_from_season_avg_window_last_5_games*)  
p-value $\beta_{1}$ : <2e-16 (t statistic 41.910)  


### Testing window sizes of 3, 5, 7, 10 days  
* No difference  
```{r, fig.height = 2,fig.width=3}
# Adjusted for player averages and window averages 
gbg_player_w_distance %>%
ggplot(aes(x = diff_sn_avg_wind_3_dist,y = diff_from_sn_avg_dist))+
  geom_point(alpha = .05)+
  geom_smooth(method="lm")+ 
  labs(y = "diff from season avg dist per game",
       x = "diff from season avg 3 day window")+
  scale_x_continuous(limits = c(-11,11))+
  scale_y_continuous(limits = c(-2.5,2.5))
#summary(lm(diff_from_sn_avg_dist ~ diff_sn_avg_wind_3_dist, data = gbg_player_w_distance))

gbg_player_w_distance %>%
ggplot(aes(x = diff_sn_avg_wind_5_dist,y = diff_from_sn_avg_dist))+
  geom_point(alpha = .05)+
  geom_smooth(method="lm")+ 
  labs(y = "diff from season avg dist per game",
       x = "diff from season avg 5 day window")+
  scale_x_continuous(limits = c(-11,11))+
  scale_y_continuous(limits = c(-2.5,2.5))

gbg_player_w_distance %>%
ggplot(aes(x = diff_sn_avg_wind_7_dist,y = diff_from_sn_avg_dist))+
  geom_point(alpha = .05)+
  geom_smooth(method="lm")+ 
  labs(y = "diff from season avg dist per game",
       x = "diff from season avg 7 day window")+
  scale_x_continuous(limits = c(-11,11))+
  scale_y_continuous(limits = c(-2.5,2.5))

gbg_player_w_distance %>%
ggplot(aes(x = diff_sn_avg_wind_10_dist,y = diff_from_sn_avg_dist))+
  geom_point(alpha = .05)+
  geom_smooth(method="lm")+ 
  labs(y = "diff from season avg dist per game",
       x = "diff from season avg 10 day window")+
  scale_x_continuous(limits = c(-11,11))+
  scale_y_continuous(limits = c(-2.5,2.5))

```

### Only players who played more than 10 minutes
* no difference in relationship  
```{r fig.height = 2,fig.width=3}
gbg_player_w_distance_play_time %>%
ggplot(aes(x = diff_sn_avg_wind_3_dist,y = diff_from_sn_avg_dist))+
  geom_point(alpha = .05)+
  geom_smooth(method="lm")+ 
  labs(y = "diff from season avg dist per game",
       x = "diff from season avg 3 day window")+
  scale_x_continuous(limits = c(-11,11))+
  scale_y_continuous(limits = c(-2.5,2.5))

gbg_player_w_distance_play_time %>%
ggplot(aes(x = diff_sn_avg_wind_5_dist,y = diff_from_sn_avg_dist))+
  geom_point(alpha = .05)+
  geom_smooth(method="lm")+ 
  labs(y = "diff from season avg dist per game",
       x = "diff from season avg 5 day window")+
  scale_x_continuous(limits = c(-11,11))+
  scale_y_continuous(limits = c(-2.5,2.5))

```

### Relationship with preformance   
* Since we showed distance to be positively correlated with fg and fg3 
percentage this small positive correlation makes sense.  

```{r fig.height = 2,fig.width=3}

gbg_player_w_distance %>%
ggplot(aes(x = diff_sn_avg_wind_7_dist, y = diff_from_sn_avg_fgpct ))+
  geom_point(alpha = .05)+
  geom_smooth(method="lm")+ 
  labs(y = "diff from season avg fg pct",
       x = "diff from season avg 7 day window")

# summary(lm(diff_from_sn_avg_fgpct ~ diff_sn_avg_wind_7_dist, data =
#           gbg_player_w_distance))

gbg_player_w_distance %>%
ggplot(aes(x = diff_sn_avg_wind_7_dist, y = diff_from_sn_avg_fg3pct ))+
  geom_point(alpha = .05)+
  geom_smooth(method="lm")+ 
  labs(y = "diff from season avg fg 3 pct",
       x = "diff from season avg 7 day window")

# summary(lm(diff_from_sn_avg_fg3pct ~ diff_sn_avg_wind_7_dist, data =
# gbg_player_w_distance))

```

1. *diff_from_season_avg_fg_pct* = $\beta_{0}$ + $\beta_{1}$ 
(*diff_sn_avg_wind_7_dist*)  
p-value $\beta_{1}$ : 1.55e-08 (t statistic 5.658)  

2. *diff_from_season_avg_fg3_pct* = $\beta_{0}$ + $\beta_{1}$
(*diff_sn_avg_wind_7_dist*)  
p-value $\beta_{1}$ : 5.42e-07 (t statistic 5.012)  


### Reltationship between travel distance and in-game distance
* Not windows just had this left over  
* No relationship  
```{r fig.height = 2,fig.width=3}
gbg_player_w_distance_subset %>%
ggplot(aes(x = distance_team_traveled_pre_game, y = diff_from_sn_avg_speed ))+
  geom_point(alpha = .05)+
  geom_smooth(method="lm")+
  labs(x = "trip distance",
       y = "diff from season avg speed")

# summary(lm(diff_from_sn_avg_speed ~ distance_team_traveled_pre_game, data =
#  gbg_player_w_distance_subset))

gbg_player_w_distance_subset %>%
ggplot(aes(x = distance_team_traveled_pre_game, y = diff_from_sn_avg_dist ))+
  geom_point(alpha = .05)+
  geom_smooth(method="lm")+
  labs(x = "trip distance",
       y = "diff from season avg distance")

# summary(lm(diff_from_sn_avg_dist ~ distance_team_traveled_pre_game, data =
#   gbg_player_w_distance_subset))

```

1. *diff_from_sn_avg_speed* = $\beta_{0}$ + $\beta_{1}$
(*distance_team_traveled_pre_game*)  
p-value $\beta_{1}$ : 0.718  (t statistic -0.362)  

2. *diff_from_sn_avg_dist* = $\beta_{0}$ + $\beta_{1}$
(*distance_team_traveled_pre_game*)  
p-value $\beta_{1}$ : 0.0215 (t statistic 2.3)  





